---
title: "GLM_WC_Pricing"
output: html_document
date: "2026-02-23"
---

# Importing Libraries and setting directories
```{r setup, include=FALSE}
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lme4)
library(glmmTMB)
library(purrr)

data_path <- '/Users/jeong-ilseung/ACTL5100/00_DataPrep/OutputData/'

wc_freq_model_clean <- read_csv(paste0(data_path, "wc_freq_model.csv"))
wc_sev_model_clean <- read_csv(paste0(data_path, "wc_sev_model.csv"))

source(paste0(getwd(), '/WC_functions.R'))
```

# Frequency Modelling
## Data check for Frequency
```{r data check freq}
source(paste0(getwd(), '/WC_freq_reconciliation.R'))
```

## Data Casting for Frequency Modelling
```{r data cast freq}
# Casting DF
df_freq <- wc_freq_model_clean %>%
  mutate(
    station_id = factor(station_id),
    solar_system = factor(solar_system),
    occupation = factor(occupation),
    employment_type = factor(employment_type),
    experience_yrs = as.numeric(experience_yrs),
    exposure = as.numeric(exposure),
    accident_history_flag = as.numeric(accident_history_flag),
    psych_stress_index = as.numeric(psych_stress_index),
    hours_per_week = as.numeric(hours_per_week),
    supervision_level = as.numeric(supervision_level),
    gravity_level = as.numeric(gravity_level),
    safety_training_index = as.numeric(safety_training_index),
    protective_gear_quality = as.numeric(protective_gear_quality),
    base_salary = as.numeric(base_salary),
    claim_count = as.numeric(claim_count)
  ) %>%
  filter(!is.na(claim_count), exposure > 0) %>%
  dplyr::select(-policy_id, - worker_id)

df_freq <- na.omit(df_freq)
```

## Poisson GLM Modelling
```{r GLM freq}

model_freq <-glm(
  claim_count ~
    solar_system +
    occupation +
    accident_history_flag +
    psych_stress_index + 
    safety_training_index +
    offset(log(exposure)),
  family = poisson(link="log"),
  data = df_freq
)

summary(model_freq)
```

# Severity Modelling
## Data check for Severity
```{r data check sev}
source(paste0(getwd(), '/WC_sev_reconciliation.R'))
```

## Data Casting for Severity Modelling
```{r data cast sev}
# Casting DF
df_sev <- wc_sev_model_clean %>%
  mutate(
    station_id = factor(station_id),
    solar_system = factor(solar_system),
    occupation = factor(occupation),
    employment_type = factor(employment_type),
    injury_type = factor(injury_type),
    injury_cause = factor(injury_cause),

    claim_length = as.numeric(claim_length),
    claim_amount = as.numeric(claim_amount),
    experience_yrs = as.numeric(experience_yrs),
    exposure = as.numeric(exposure),
    accident_history_flag = as.numeric(accident_history_flag),
    psych_stress_index = as.numeric(psych_stress_index),
    hours_per_week = as.numeric(hours_per_week),
    supervision_level = as.numeric(supervision_level),
    gravity_level = as.numeric(gravity_level),
    safety_training_index = as.numeric(safety_training_index),
    protective_gear_quality = as.numeric(protective_gear_quality),
    base_salary = as.numeric(base_salary),
    log_salary = log(base_salary)
  ) %>%
  filter(
    !is.na(claim_amount),
    claim_amount > 0,
    !is.na(base_salary),
    base_salary > 0
  ) %>%
  dplyr::select(-claim_id, -claim_seq, -policy_id, -worker_id, -claim_length,-claim_count)

df_sev <- na.omit(df_sev)
```

## Gamma GLM Modelling
```{r GLM sev}

model_sev <- glm(
  claim_amount ~
    psych_stress_index +
    hours_per_week +
    supervision_level +
    gravity_level +
    safety_training_index +
    protective_gear_quality +
    log_salary +
    employment_type +
    occupation +
    solar_system,
  family = Gamma(link = "log"),
  data = df_sev
)

summary(model_sev)
```

# Probability Table for Injury Cause and Type by Occupation
```{r Injury Probability Model}
# Injury probability table
injury_prob_table <- df_sev %>%
  group_by(occupation, injury_type, injury_cause) %>%
  summarise(n = n(), .groups = "drop") %>%
  group_by(occupation) %>%
  mutate(prob = n / sum(n)) %>%
  ungroup()

# fallback table if an occupation has no matching injury profile
injury_prob_table_all <- df_sev %>%
  group_by(injury_type, injury_cause) %>%
  summarise(n = n(), .groups = "drop") %>%
  mutate(prob = n / sum(n))
```

# NB Table
```{r NB generator}
# Note: This process is computationally intensive. 
# If regeneration is not required, consider using the saved dataset instead.
# Random Seed set to 123
set.seed(123)

# NB config file
source(paste0(getwd(), '/WC_NB_mapping_table.R'))

# Create portfolio
nb_data_base <- make_nb_wc_data(
  df_freq = df_freq,
  df_sev = df_sev,
  nb_workforce = nb_workforce,
  occupation_map = occupation_map
)

nb_10y <- make_nb_wc_data_10y(
  df_freq        = df_freq,
  df_sev         = df_sev,
  nb_workforce   = nb_workforce,
  occupation_map = occupation_map,
  worker_influx  = worker_influx
)

nb_entrants <- nb_10y$entrants
nb_panel    <- nb_10y$panel

# saveRDS(nb_entrants, file = paste0(getwd(), "/nb_entrants.rds"))
# saveRDS(nb_panel, file = paste0(getwd(), "/nb_panel.rds"))

# nb_entrants <- readRDS(paste0(getwd(), '/nb_entrants.rds'))
# nb_panel <- readRDS(paste0(getwd(), '/nb_panel.rds'))
```

# NB Pricing
```{r NB Pricing}
# Filter initial year NB only for short term
nb_data_base <- nb_entrants %>% filter(added_year <= 1)
# Expected Loss Calculation
# Note: This process is computationally intensive. 
# If regeneration is not required, consider using the saved dataset instead.
source(paste0(getwd(), '/WC_NB_data_generator.R'))
# expected_sev_by_worker_base <- readRDS(paste0(getwd(), '/expected_sev_by_worker_base.rds'))

# Parameter settings
loading <- 0.20 # Premium Loading to the expected loss

# Compute frequency prediction
freq_pred_base <- predict(
  model_freq,
  newdata = nb_data_base,
  type = "response"
)

# Calculate aggregate premium
expected_loss_by_worker_base <- freq_pred_base * expected_sev_by_worker_base
aggregate_expected_loss_base <- sum(expected_loss_by_worker_base, na.rm = TRUE)
aggregate_earnings_base      <- aggregate_expected_loss_base * (1 + loading)
```

# NB Short-term Simulation
```{r NB short-term simulation}
# Run simulations
# Note: This process is computationally intensive. 
# If regeneration is not required, consider using the saved dataset instead.

n_sim <- 10000
sim_list <- vector("list", n_sim)

# Sigma of Gamma
phi_sev <- sigma(model_sev)

for (sim in seq_len(n_sim)) {
  sim_list[[sim]] <- simulate_one_nb_wc_fast(
    nb_data_base = nb_data_base,
    freq_pred_base = freq_pred_base,
    aggregate_expected_loss_base = aggregate_expected_loss_base,
    aggregate_earnings_base = aggregate_earnings_base,
    model_sev = model_sev,
    df_sev = df_sev,
    injury_prob_table = injury_prob_table,
    injury_prob_table_all = injury_prob_table_all,
    phi_sev = phi_sev
  )
  
  message("Completed simulation iteration: ", sim, " / ", n_sim)
}

sim_results_wc <- bind_rows(sim_list)

# saveRDS(sim_results_wc, file = paste0(getwd(), "/sim_results_wc_s2.rds"))
# sim_results_wc <- readRDS(paste0(getwd(), '/sim_results_wc.rds'))

# Summary
summary(sim_results_wc)

## Aggregate Loss
mean(sim_results_wc$aggregate_loss)
sd(sim_results_wc$aggregate_loss)
quantile(sim_results_wc$aggregate_loss, c(0.5, 0.9, 0.95, 0.99, 0.995))

# Probability of Loss
mean(sim_results_wc$profit < 0)

## Aggregate Loss
mean(sim_results_wc$profit)
sd(sim_results_wc$profit)
quantile(sim_results_wc$profit, c(0.005, 0.01,0.05, 0.1, 0.5))

# Plots
hist(
  sim_results_wc$aggregate_loss,
  breaks = 40,
  main = "Aggregate Loss Distribution for NB Workers' Comp Portfolio",
  xlab = "Aggregate Loss"
)

hist(
  sim_results_wc$profit,
  breaks = 40,
  main = "Profit Distribution for NB Workers' Comp Portfolio",
  xlab = "Profit"
)
```


# NB Long-term Simulation
```{r Long-term Simulation}
# Filter full entrant table once
nb_data_all <- nb_entrants

# Expected Loss Calculation
# Note: This process is computationally intensive.
# If regeneration is not required, consider using the saved dataset instead.
source(paste0(getwd(), '/WC_NB_data_generator_all.R'))
# expected_sev_by_worker_base <- readRDS(paste0(getwd(), '/expected_sev_by_worker_base_all.rds'))

freq_pred_all <- predict(
  model_freq,
  newdata = nb_data_all,
  type = "response",
  re.form = NA,
  allow.new.levels = TRUE
)

# expected_sev_by_worker_base should correspond to nb_data_all
# If not, you need the full vector saved for all entrants, not just year 1
expected_sev_all <- expected_sev_by_worker_base
expected_loss_by_worker_all <- freq_pred_all * expected_sev_all

inflation_vec <- rep(0.025, 10)
interest_vec  <- rep(0.018, 10)

loading <- 0.20
phi_sev <- sigma(model_sev)

n_sim <- 100000
sim_list_10yr <- vector("list", n_sim)

for (sim in seq_len(n_sim)) {
  sim_list_10yr[[sim]] <- simulate_10yr_nb_wc(
    nb_data_all = nb_entrants,
    freq_pred_all = freq_pred_all,
    expected_loss_by_worker_all = expected_loss_by_worker_all,
    model_sev = model_sev,
    df_sev = df_sev,
    injury_prob_table = injury_prob_table,
    injury_prob_table_all = injury_prob_table_all,
    phi_sev = phi_sev,
    loading = loading,
    inflation_vec = inflation_vec,
    interest_vec = interest_vec
  )

  message("Completed 10-year simulation iteration: ", sim, " / ", n_sim)
}

sim_results_10yr <- bind_rows(sim_list_10yr)

# saveRDS(sim_results_10yr, file = paste0(getwd(), "/sim_results_10yr.rds"))
# sim_results_10yr <- readRDS(paste0(getwd(), '/sim_results_10yr.rds'))


summary(sim_results_10yr)

mean(sim_results_10yr$pv_aggregate_loss)
sd(sim_results_10yr$pv_aggregate_loss)
quantile(sim_results_10yr$pv_aggregate_loss, c(0.5, 0.9, 0.95, 0.99, 0.995))

mean(sim_results_10yr$pv_aggregate_profit)
sd(sim_results_10yr$pv_aggregate_profit)
quantile(sim_results_10yr$pv_aggregate_profit, c(0.005, 0.01, 0.05, 0.1, 0.5))

mean(sim_results_10yr$pv_aggregate_profit < 0)

hist(
  sim_results_10yr$pv_aggregate_loss,
  breaks = 40,
  main = "PV Aggregate Loss Distribution (10-Year WC NB Portfolio)",
  xlab = "PV Aggregate Loss"
)

hist(
  sim_results_10yr$pv_aggregate_profit,
  breaks = 40,
  main = "PV Aggregate Profit Distribution (10-Year WC NB Portfolio)",
  xlab = "PV Aggregate Profit"
)
```