---
title: "GLM_CARGO_claim"
output: html_document
date: "2026-03-14"
---

```{r setup, include=FALSE}
install.packages("lme4")
install.packages("glmmTMB")
install.packages("kableExtra")

library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lme4)
library(glmmTMB)
library(kableExtra)
library(MASS)

bronze_path <- paste0(getwd(), '/Bronze/')

cargo_freq_model_clean <- read_csv(paste0(bronze_path, "cargo_claim_freq_update1.csv"))
cargo_sev_model_clean <- read_csv(paste0(bronze_path, "cargo_claim_sev_update1.csv"))
```

# CHECK DATA CLEANING
```{r}
######## Range Checks ########
cat("=== FREQ TABLE RANGE CHECKS ===\n")
# 1. claim_count (valid: 0-5)
cat("claim_count:", min(cargo_freq_model_clean$claim_count), "to", max(cargo_freq_model_clean$claim_count), "\n")
# 2. cargo_value (valid: 50K-680,000K)
cat("cargo_value:", format(min(cargo_freq_model_clean$cargo_value), scientific = FALSE), "to", format(max(cargo_freq_model_clean$cargo_value), scientific = FALSE), "\n")
# 3. weight (valid: 1,500-250,000)
cat("weight:", min(cargo_freq_model_clean$weight), "to", max(cargo_freq_model_clean$weight), "\n")
# 4. route_risk (valid: 1-5)
cat("route_risk unique values:", unique(sort(cargo_freq_model_clean$route_risk)), "\n")
# 5. distance (valid: 1-100)
cat("distance:", min(cargo_freq_model_clean$distance), "to", max(cargo_freq_model_clean$distance), "\n")
# 6. transit_duration (valid: 1-60)
cat("transit_duration:", min(cargo_freq_model_clean$transit_duration), "to", max(cargo_freq_model_clean$transit_duration), "\n")
# 7. pilot_experience (valid: 1-30)
cat("pilot_experience:", min(cargo_freq_model_clean$pilot_experience), "to", max(cargo_freq_model_clean$pilot_experience), "\n")
# 8. vessel_age (valid: 1-50)
cat("vessel_age:", min(cargo_freq_model_clean$vessel_age), "to", max(cargo_freq_model_clean$vessel_age), "\n")
# 9. solar_radiation (valid: 0-1)
cat("solar_radiation:", min(cargo_freq_model_clean$solar_radiation), "to", max(cargo_freq_model_clean$solar_radiation), "\n")
# 10. debris_density (valid: 0-1)
cat("debris_density:", min(cargo_freq_model_clean$debris_density), "to", max(cargo_freq_model_clean$debris_density), "\n")
# 11. exposure (valid: 0-1)
cat("exposure:", min(cargo_freq_model_clean$exposure), "to", max(cargo_freq_model_clean$exposure), "\n")
# 12. cargo_type (categorical)
cat("cargo_type unique values:", unique(cargo_freq_model_clean$cargo_type), "\n")
# 13. container_type (categorical)
cat("container_type unique values:", unique(cargo_freq_model_clean$container_type), "\n")
# Check a few raw values
head(cargo_freq_model_clean$cargo_type, 20)

cat("\n=== SEV TABLE RANGE CHECKS ===\n")
# 1. claim_amount
cat("claim_amount:", format(min(cargo_sev_model_clean$claim_amount), scientific = FALSE), "to", format(max(cargo_sev_model_clean$claim_amount), scientific = FALSE), "\n")
# 2. cargo_value
cat("cargo_value:", format(min(cargo_sev_model_clean$cargo_value), scientific = FALSE), "to", format(max(cargo_sev_model_clean$cargo_value), scientific = FALSE), "\n")
# 3. weight (valid: 1,500-250,000)
cat("weight:", min(cargo_sev_model_clean$weight), "to", max(cargo_sev_model_clean$weight), "\n")
# 4. route_risk (valid: 1-5)
cat("route_risk unique values:", unique(sort(cargo_sev_model_clean$route_risk)), "\n")
# 5. distance (valid: 1-100)
cat("distance:", min(cargo_sev_model_clean$distance), "to", max(cargo_sev_model_clean$distance), "\n")
# 6. transit_duration (valid: 1-60)
cat("transit_duration:", min(cargo_sev_model_clean$transit_duration), "to", max(cargo_sev_model_clean$transit_duration), "\n")
# 7. pilot_experience (valid: 1-30)
cat("pilot_experience:", min(cargo_sev_model_clean$pilot_experience), "to", max(cargo_sev_model_clean$pilot_experience), "\n")
# 8. vessel_age (valid: 1-50)
cat("vessel_age:", min(cargo_sev_model_clean$vessel_age), "to", max(cargo_sev_model_clean$vessel_age), "\n")
# 9. solar_radiation (valid: 0-1)
cat("solar_radiation:", min(cargo_sev_model_clean$solar_radiation), "to", max(cargo_sev_model_clean$solar_radiation), "\n")
# 10. debris_density (valid: 0-1)
cat("debris_density:", min(cargo_sev_model_clean$debris_density), "to", max(cargo_sev_model_clean$debris_density), "\n")
# 11. exposure (valid: 0-1)
cat("exposure:", min(cargo_sev_model_clean$exposure), "to", max(cargo_sev_model_clean$exposure), "\n")
# 12. cargo_type (categorical)
cat("cargo_type unique values:", unique(cargo_sev_model_clean$cargo_type), "\n")
# 13. container_type (categorical)
cat("container_type unique values:", unique(cargo_sev_model_clean$container_type), "\n")
```

# DATA FORMATING
```{r}
# Casting DF — Freq table
df_cargo_freq <- cargo_freq_model_clean %>%
  mutate(
    cargo_value      = as.numeric(cargo_value),
    weight           = as.numeric(weight),
    distance         = as.numeric(distance),
    transit_duration = as.numeric(transit_duration),
    pilot_experience = as.numeric(pilot_experience),
    vessel_age       = as.numeric(vessel_age),
    solar_radiation  = as.numeric(solar_radiation),
    debris_density   = as.numeric(debris_density),
    exposure         = as.numeric(exposure),
    route_risk       = factor(route_risk, levels = 1:5, ordered = TRUE),
    cargo_type       = factor(cargo_type),
    container_type   = factor(container_type)
  ) %>%
  filter(!is.na(claim_count), exposure > 0)

# Casting DF — Sev table
df_cargo_sev <- cargo_sev_model_clean %>%
  mutate(
    cargo_value      = as.numeric(cargo_value),
    weight           = as.numeric(weight),
    distance         = as.numeric(distance),
    transit_duration = as.numeric(transit_duration),
    pilot_experience = as.numeric(pilot_experience),
    vessel_age       = as.numeric(vessel_age),
    solar_radiation  = as.numeric(solar_radiation),
    debris_density   = as.numeric(debris_density),
    exposure         = as.numeric(exposure),
    route_risk       = factor(route_risk, levels = 1:5, ordered = TRUE),
    cargo_type       = factor(cargo_type),
    container_type   = factor(container_type)
  ) %>%
  filter(!is.na(claim_amount), claim_amount > 0)

# Verify
nrow(df_cargo_freq)
nrow(df_cargo_sev)
glimpse(df_cargo_freq)
```

# Exploratory Data Analysis 1
```{r}
cat("Mean of claim_count:", mean(df_cargo_freq$claim_count), "\n")
cat("Variance of claim_count:", var(df_cargo_freq$claim_count), "\n")
cat("Variance/Mean ratio:", var(df_cargo_freq$claim_count) / mean(df_cargo_freq$claim_count), "\n")
```

# Exploratory Data Analysis 2
```{r}
summary(df_cargo_freq$claim_count)
barplot(table(df_cargo_freq$claim_count),
        main = "Claim Count Distribution",
        xlab = "Claim Count",
        ylab = "Number of Policies",
        col  = "steelblue")

summary(df_cargo_sev$claim_amount)
# Raw claim amount
ggplot(df_cargo_sev, aes(x = claim_amount)) +
  geom_histogram(bins = 50, fill = "steelblue") +
  labs(title = "Distribution of Claim Amount")

# Log transformed
ggplot(df_cargo_sev, aes(x = log(claim_amount))) +
  geom_histogram(bins = 50, fill = "steelblue") +
  labs(title = "Distribution of Log Claim Amount")

# See if cargo_type explains the bimodality
ggplot(df_cargo_sev, aes(x = log(claim_amount), fill = cargo_type)) +
  geom_histogram(bins = 50, alpha = 0.6) +
  labs(title = "Log Claim Amount by Cargo Type")
```

# CHECK CARGO & CONTAINER FIT FREQ MODELLING 
```{r}
# Test for caro and container type (categorical variable)
# ─ Base model (neither cargo_type nor container_type) ─
model_base <- glm(
  claim_count ~ cargo_value + weight + distance + 
    transit_duration + pilot_experience + vessel_age + 
    solar_radiation + debris_density + route_risk + 
    offset(log(exposure)),
  family = poisson(link = "log"),
  data = df_cargo_freq)

# ─ Model with cargo_type only ─
model_with_cargo <- glm(
  claim_count ~ cargo_value + weight + distance + 
    transit_duration + pilot_experience + vessel_age + 
    solar_radiation + debris_density + route_risk + 
    cargo_type +
    offset(log(exposure)),
  family = poisson(link = "log"),
  data = df_cargo_freq)

# ─ Model with container_type only ──
model_with_container <- glm(
  claim_count ~ cargo_value + weight + distance + 
    transit_duration + pilot_experience + vessel_age + 
    solar_radiation + debris_density + route_risk + 
    container_type +
    offset(log(exposure)),
  family = poisson(link = "log"),
  data = df_cargo_freq)

# ─ Model with both ─
model_with_both <- glm(
  claim_count ~ cargo_value + weight + distance + 
    transit_duration + pilot_experience + vessel_age + 
    solar_radiation + debris_density + route_risk + 
    cargo_type + container_type +
    offset(log(exposure)),
  family = poisson(link = "log"),
  data = df_cargo_freq)

# ─ Test 1: Does cargo_type add value on top of container_type? ─
cat("=== Testing cargo_type (controlling for container_type) ===\n")
anova(model_with_container, model_with_both, test = "Chisq")

# ─ Test 2: Does container_type add value on top of cargo_type? ─
cat("=== Testing container_type (controlling for cargo_type) ===\n")
anova(model_with_cargo, model_with_both, test = "Chisq")

# both cargo and container types are significant to add in.
```

# FREQ GLM FIT - POISSON
```{r}
# Model 1a: Frequency Model (Poisson GLM) 
model_cargo_freq1 <- glm(
  claim_count ~ cargo_value + weight + distance + transit_duration + pilot_experience + vessel_age + solar_radiation + debris_density + route_risk + cargo_type + container_type + offset(log(exposure)),
  family = poisson(link = "log"),
  data = df_cargo_freq
)

summary(model_cargo_freq1)

```
GLM equation: log(E[claim_count])=fixed effects+log(exposure)

# POISSON DISPERSION CHECK
```{r}
# ratio check poisson suitable? 
deviance(model_cargo_freq1) / df.residual(model_cargo_freq1)

# Refit with Quasi Poisson
model_freq_quasi <- glm(claim_count ~ cargo_value + weight + distance + transit_duration + pilot_experience + vessel_age + solar_radiation + debris_density + route_risk + cargo_type + container_type + offset(log(exposure)),
                        family = quasipoisson(link = "log"),
                        data = df_cargo_freq)
summary(model_freq_quasi)
```

# FREQ GLM FIT: NEGATIVE BINOMIAL
```{r}
# Model 2: Fit Negative Binomial 
model_cargo_freq2 <- glm.nb(
  claim_count ~ cargo_value + weight + distance + transit_duration + pilot_experience + vessel_age + solar_radiation + debris_density + route_risk + cargo_type + container_type + offset(log(exposure)),
  data = df_cargo_freq)

summary(model_cargo_freq2)
```

# NB DISPERSION CHECK
```{r}
# Dispersion ratio = residual deviance / degrees of freedom
dispersion_ratio <- model_cargo_freq2$deviance / 
                    model_cargo_freq2$df.residual

cat("Dispersion ratio:", dispersion_ratio, "\n")

```

# TRIMMED NB - FINAL 
```{r}
model_cargo_freq3 <- glm.nb(claim_count ~ cargo_value + pilot_experience + route_risk + cargo_type + container_type + offset(log(exposure)),
  data = df_cargo_freq)

summary(model_cargo_freq3)

# Compare AIC NB
cat("Full NB AIC:", AIC(model_cargo_freq2), "\n")
cat("Trimmed NB AIC:", AIC(model_cargo_freq3), "\n")
```

Final freq model :model_cargo_freq3


# SEV GLM FIT: GAMMA 
```{r}
model_cargo_sev1 <- glm( claim_amount ~ cargo_value + weight + distance + transit_duration + pilot_experience + vessel_age + solar_radiation + debris_density + route_risk + cargo_type + container_type,
  family = Gamma(link = "log"), data = df_cargo_sev)

summary(model_cargo_sev1)
```

# TRIMMED SEV GLM FIT: GAMMA 
```{r}
# Trimmed severity model 
model_cargo_sev2 <- glm( claim_amount ~ cargo_value + weight + solar_radiation +  debris_density + route_risk + cargo_type,
  family = Gamma(link = "log"),
  data = df_cargo_sev)

summary(model_cargo_sev2)

# Compare AIC GAMMA
cat("Full Gamma AIC:", AIC(model_cargo_sev1), "\n")
cat("Trimmed Gamma AIC:", AIC(model_cargo_sev2), "\n")
```
Final sev model :model_cargo_sev2

# TRIMMED GAMMA MODEL RESIDUAL PLOT 
```{r}
par(mfrow = c(2, 2))
plot(model_cargo_sev2)
```

# SEV FITTED MODEL CARGO TYPE DISTRIBUTION 
```{r}
boxplot(fitted(model_cargo_sev2) ~ df_cargo_sev$cargo_type,
        las = 2, main = "Fitted values by cargo type")
```
The severity model shows evidence of heavy-tailed claim behaviour in extreme scenarios. 

# NB INVENTORY DATA
```{r}
set.seed(123)

# Fleet sizes from inventory data
n_helionis <- 1160
n_bayesia  <- 1128
n_oryn     <- 774
n_total    <- n_helionis + n_bayesia + n_oryn  # 3062
```

# SAMPLING NB 
```{r}
make_nb_data_cargo <- function(df_freq, df_sev, n_helionis, n_bayesia, n_oryn) {
  
  # Container counts per solar system
  container_helionis <- c(
    rep("DeepSpace Haulbox",       58),
    rep("DockArc Freight Case",   116),
    rep("HardSeal Transit Crate", 580),
    rep("LongHaul Vault Canister",232),
    rep("QuantumCrate Module",    174)
  )
  
  container_bayesia <- c(
    rep("DeepSpace Haulbox",       56),
    rep("DockArc Freight Case",   113),
    rep("HardSeal Transit Crate", 564),
    rep("LongHaul Vault Canister",226),
    rep("QuantumCrate Module",    169)
  )
  
  container_oryn <- c(
    rep("DeepSpace Haulbox",       39),
    rep("DockArc Freight Case",    77),
    rep("HardSeal Transit Crate", 387),
    rep("LongHaul Vault Canister",155),
    rep("QuantumCrate Module",    116)
  )
  
  total_n <- n_helionis + n_bayesia + n_oryn
  
  sampled_rows <- df_sev[sample(seq_len(nrow(df_sev)), total_n, replace = TRUE), ] #row based sampling to keep the feature to the row
  
  nb_data <- data.frame(
    cargo_value      = sampled_rows$cargo_value,
    pilot_experience = sampled_rows$pilot_experience,
    route_risk       = sampled_rows$route_risk,
    cargo_type       = sampled_rows$cargo_type,
    weight           = sampled_rows$weight,
    solar_radiation  = sampled_rows$solar_radiation,
    debris_density   = sampled_rows$debris_density,
    exposure         = 1
  )
  
  # Replace with required future portfolio container mix
  nb_data$container_type <- factor(
    c(container_helionis, container_bayesia, container_oryn),
    levels = levels(df_freq$container_type)
  )
  
  # Solar system tags
  nb_data$solar_system <- c(
    rep("Helionis Cluster", n_helionis),
    rep("Bayesia System",   n_bayesia),
    rep("Oryn Delta",       n_oryn)
  )
  
  # Match factor levels
  nb_data$cargo_type <- factor(nb_data$cargo_type, 
                               levels = levels(df_freq$cargo_type))
  nb_data$route_risk <- factor(nb_data$route_risk,
                               levels = levels(df_freq$route_risk),
                               ordered = TRUE)
  
  # Cap weight
  container_max_weight <- c(
    "DeepSpace Haulbox"       = 25000,
    "DockArc Freight Case"    = 50000,
    "HardSeal Transit Crate"  = 100000,
    "LongHaul Vault Canister" = 150000,
    "QuantumCrate Module"     = 250000
  )
  
  nb_data <- nb_data %>%
    mutate(
      weight = pmin(weight,
                    container_max_weight[as.character(container_type)]))
  
  # Verify all set
  cat("Weight cap verification:\n")
  print(nb_data %>%
    group_by(container_type) %>%
    summarise(
      max_weight  = max(weight),
      mean_weight = round(mean(weight), 0),
      .groups = "drop"
    ))
  
  return(nb_data)
}
```

# GENERARE NB PORTFOLIO
```{r}
nb_data_cargo <- make_nb_data_cargo(
  df_freq    = df_cargo_freq,
  df_sev     = df_cargo_sev,
  n_helionis = n_helionis,
  n_bayesia  = n_bayesia,
  n_oryn     = n_oryn
)

glimpse(nb_data_cargo)
table(nb_data_cargo$solar_system)
table(nb_data_cargo$container_type)
table(nb_data_cargo$cargo_type)
```

# CHECK NB PORTFOLIO - all criteria satisfied 
```{r}
weight_check <- nb_data_cargo %>%
  group_by(solar_system, container_type) %>%
  summarise(
    n_vessels      = n(),
    max_weight     = max(weight),
    total_capacity = max(weight) * n(),
    actual_total   = sum(weight),
    is_exceeded    = sum(weight) > max(weight) * n(),
    .groups = "drop"
  ) %>%
  arrange(solar_system, container_type)

print(weight_check)
```
# NB DATA DISTRIBUTION
```{r}
nb_data_cargo %>%
  group_by(cargo_type) %>%
  summarise(
    mean_weight = round(mean(weight), 0),
    max_weight  = max(weight)
  ) %>%
  arrange(desc(mean_weight))

nb_data_cargo %>%
  group_by(cargo_type) %>%
  summarise(
    mean_weight      = round(mean(weight), 0),
    mean_cargo_value = round(mean(cargo_value), 0)
  ) %>%
  arrange(desc(mean_cargo_value))
```


# QUICK NB DATA PREDICTION 
```{r}
# Predict claim frequency
nb_data_cargo$freq_pred <- predict(
  model_cargo_freq3,
  newdata = nb_data_cargo,
  type = "response"
)

# Predict claim severity
nb_data_cargo$sev_pred <- predict(
  model_cargo_sev2,
  newdata = nb_data_cargo,
  type = "response"
)

# Expected loss per unit
nb_data_cargo$expected_loss <- nb_data_cargo$freq_pred * nb_data_cargo$sev_pred

cat("Mean predicted claim frequency:", mean(nb_data_cargo$freq_pred), "\n")
cat("Mean predicted claim severity:", mean(nb_data_cargo$sev_pred), "\n")
cat("Mean predicted loss per policy:", mean(nb_data_cargo$expected_loss), "\n")
cat("Total expected loss:", sum(nb_data_cargo$expected_loss), "\n")

aggregate(expected_loss ~ cargo_type, data = nb_data_cargo, sum)
aggregate(expected_loss ~ container_type, data = nb_data_cargo, sum)
aggregate(sev_pred ~ cargo_type, data = nb_data_cargo, mean)
```

# SPLIT CARGO TYPE SEGMENT 
```{r}
nb_data_cargo$segment <- ifelse(
  nb_data_cargo$cargo_type %in% c("gold", "platinum"),
  "high_value",
  "standard"
)

table(nb_data_cargo$segment)
aggregate(expected_loss ~ segment, data = nb_data_cargo, sum)
```

# PARAMETERS
```{r}
nb_high <- subset(nb_data_cargo, segment == "high_value")
nb_std  <- subset(nb_data_cargo, segment == "standard")

theta <- model_cargo_freq3$theta
gamma_disp <- summary(model_cargo_sev2)$dispersion
```

# SIMULATION FUNCTION
```{r}
simulate_aggregate_loss <- function(freq_pred, sev_pred, theta, gamma_disp, n_sim) {
  
  n_units <- length(freq_pred)
  agg_loss <- numeric(n_sim)
  gamma_shape <- 1 / gamma_disp
  
  for (i in 1:n_sim) {
    
    claim_counts <- rnbinom(n_units, mu = freq_pred, size = theta)
    total_loss <- 0
    
    for (j in 1:n_units) {
      if (claim_counts[j] > 0) {
        gamma_rate <- gamma_shape / sev_pred[j]
        claims <- rgamma(claim_counts[j], shape = gamma_shape, rate = gamma_rate)
        total_loss <- total_loss + sum(claims)
      }
    }
    
    agg_loss[i] <- total_loss
  }
  
  return(agg_loss)
}
```


# 100000 SIMULATION, SETSEED(123)
```{r}
set.seed(123)
n_sim <- 100000

cat("Running simulation for high-value segment...\n")
agg_high <- simulate_aggregate_loss(
  freq_pred = nb_high$freq_pred,
  sev_pred = nb_high$sev_pred,
  theta = theta,
  gamma_disp = gamma_disp,
  n_sim = n_sim
)

cat("Running simulation for standard segment...\n")
agg_std <- simulate_aggregate_loss(
  freq_pred = nb_std$freq_pred,
  sev_pred = nb_std$sev_pred,
  theta = theta,
  gamma_disp = gamma_disp,
  n_sim = n_sim
)

# Combined portfolio
agg_total <- agg_high + agg_std
```

# SIMULATION RESULT 
```{r}
summarise_loss <- function(x, n_policies) {
  c(
    
    riskcount = n_policies,
    Loss_Mean       = mean(x),
    VaR50      = quantile(x, 0.50),
    VaR90      = quantile(x, 0.90),
    VaR95      = quantile(x, 0.95),
    VaR99      = quantile(x, 0.99),
    VaR99.95    = quantile(x, 0.9995),
    SD         = sd(x),
    Min        = min(x),
    Median     = quantile(x, 0.50),
    TVaR95     = mean(x[x > quantile(x, 0.95)]),
    TVaR99     = mean(x[x > quantile(x, 0.99)]),
    TVaR99.5   = mean(x[x > quantile(x, 0.9995)])
  )
}

summary_high  <- summarise_loss(agg_high,  n_policies = nrow(nb_high))
summary_std   <- summarise_loss(agg_std,   n_policies = nrow(nb_std))
summary_total <- summarise_loss(agg_total, n_policies = nrow(nb_data_cargo))

summary_high
summary_std
summary_total
```

# SIMULATION RESULT DISTRIBUTION
```{r}
par(mfrow = c(1, 3))

hist(agg_high, breaks = 50, main = "High-value Segment",
     xlab = "Aggregate Loss", col = "lightcoral", border = "white")

hist(agg_std, breaks = 50, main = "Standard Segment",
     xlab = "Aggregate Loss", col = "lightblue", border = "white")

hist(agg_total, breaks = 50, main = "Total Portfolio",
     xlab = "Aggregate Loss", col = "lightgreen", border = "white")
```

# LOADING SIMULATION - PARAMETERS AND SEGMENTS
```{r}
set.seed(123)

# Parameters 
loading <- 0.40
n_sim   <- 100000

# Segment flag 
nb_data_cargo$segment <- ifelse(
  nb_data_cargo$cargo_type %in% c("gold", "platinum"),
  "high_value",
  "standard")

nb_high <- subset(nb_data_cargo, segment == "high_value")
nb_std  <- subset(nb_data_cargo, segment == "standard")

# Model parameters 
theta      <- model_cargo_freq3$theta
gamma_disp <- summary(model_cargo_sev2)$dispersion
gamma_shape <- 1 / gamma_disp
```

# SIMULATION FUNCTION
```{r}
simulate_segment <- function(freq_pred, sev_pred, theta, gamma_shape, loading, n_sim) {
  
  n_units <- length(freq_pred)
  
  # Expected premium revenue based on loaded expected loss
  expected_loss_by_policy <- freq_pred * sev_pred
  premium_revenue <- sum(expected_loss_by_policy * (1 + loading))
  
  results <- data.frame(
    aggregate_cost   = numeric(n_sim),   # claims cost
    revenue          = numeric(n_sim),   # premium revenue
    net_revenue      = numeric(n_sim),   # underwriting profit
    loss_ratio       = numeric(n_sim)
  )
  
  for (i in 1:n_sim) {
    
    claim_counts <- rnbinom(n_units, mu = freq_pred, size = theta)
    total_loss <- 0
    
    for (j in 1:n_units) {
      if (claim_counts[j] > 0) {
        gamma_rate <- gamma_shape / sev_pred[j]
        claims <- rgamma(claim_counts[j], shape = gamma_shape, rate = gamma_rate)
        total_loss <- total_loss + sum(claims)
      }
    }
    
    results$aggregate_cost[i] <- total_loss
    results$revenue[i]        <- premium_revenue
    results$net_revenue[i]    <- premium_revenue - total_loss
    results$loss_ratio[i]     <- total_loss / premium_revenue
  }
  
  return(results)
}
```

# RUN SIMULATION 
```{r}
cat("Running segmented simulation...\n")

sim_high <- simulate_segment(
  freq_pred   = nb_high$freq_pred,
  sev_pred    = nb_high$sev_pred,
  theta       = theta,
  gamma_shape = gamma_shape,
  loading     = loading,
  n_sim       = n_sim
)

sim_std <- simulate_segment(
  freq_pred   = nb_std$freq_pred,
  sev_pred    = nb_std$sev_pred,
  theta       = theta,
  gamma_shape = gamma_shape,
  loading     = loading,
  n_sim       = n_sim
)
```

# AGGREGATE SIMULATION RESULT
```{r}
sim_total <- data.frame(
  aggregate_cost = sim_high$aggregate_cost + sim_std$aggregate_cost,
  revenue        = sim_high$revenue + sim_std$revenue,
  net_revenue    = sim_high$net_revenue + sim_std$net_revenue,
  loss_ratio     = (sim_high$aggregate_cost + sim_std$aggregate_cost) /
                   (sim_high$revenue + sim_std$revenue))
```

# SIMULATION RESULT SUMMARY
```{r}
summarise_results <- function(sim_df, n_policies, loading) {
  data.frame(
    Risk_count = n_policies,
    loading = paste0(loading * 100, "%"),
    Premium = mean(sim_df$revenue),
    
    Mean_Cost        = mean(sim_df$aggregate_cost),
    SD_Cost          = sd(sim_df$aggregate_cost),
    VAR50_Cost             = quantile(sim_df$aggregate_cost, 0.50),
    VAR90_Cost             = quantile(sim_df$aggregate_cost, 0.90),
    VAR95_Cost             = quantile(sim_df$aggregate_cost, 0.95),
    VAR99_Cost             = quantile(sim_df$aggregate_cost, 0.99),
    VAR99.95_Cost           = quantile(sim_df$aggregate_cost, 0.9995),
    TVaR95_Cost      = mean(sim_df$aggregate_cost[sim_df$aggregate_cost > quantile(sim_df$aggregate_cost, 0.95)]),
    TVaR99_Cost      = mean(sim_df$aggregate_cost[sim_df$aggregate_cost > quantile(sim_df$aggregate_cost, 0.99)]),
    TVaR99.95_Cost    = mean(sim_df$aggregate_cost[sim_df$aggregate_cost > quantile(sim_df$aggregate_cost, 0.9995)]),
    
    Mean_NetRevenue  = mean(sim_df$net_revenue),
    SD_NetRevenue    = sd(sim_df$net_revenue),
    Prob_Loss        = mean(sim_df$net_revenue < 0),
    VaR50_NetRevenue = quantile(sim_df$net_revenue, 0.50),
    VaR90_NetRevenue = quantile(sim_df$net_revenue, 0.10),
    VaR95_NetRevenue = quantile(sim_df$net_revenue, 0.05),
    VaR99_NetRevenue = quantile(sim_df$net_revenue, 0.01),
    VaR99.95_NetRevenue = quantile(sim_df$net_revenue, 0.0005),
    
    Mean_LossRatio   = mean(sim_df$loss_ratio),
    Median_LossRatio = median(sim_df$loss_ratio)
  )
}
```

# SIMULATION RESULT SUMMARY
```{r}
# calls
summary_high  <- summarise_results(sim_high,  n_policies = nrow(nb_high),        loading = loading)
summary_std   <- summarise_results(sim_std,   n_policies = nrow(nb_std),         loading = loading)
summary_total <- summarise_results(sim_total, n_policies = nrow(nb_data_cargo),  loading = loading)

cat("\n=== High Value Segment ===\n")
print(summary_high)

cat("\n=== Standard Segment ===\n")
print(summary_std)

cat("\n=== Total Portfolio ===\n")
print(summary_total)
```

# SUMMARY OUTPUT
```{r}

cat("\n=== SHORT TERM CARGO SIMULATION RESULTS ===\n")
cat("Risk Count:  ", nrow(nb_data_cargo), "\n")
cat("Loading:      40%\n")

# Premium 
cat("\n--- Premium (Fixed) ---\n")
cat("Aggregate Premium: ", round(mean(sim_total$revenue), 0), "\n")

# Aggregate Loss 
cat("\n--- Aggregate Loss ---\n")
cat("Mean Aggregate Loss: ", round(mean(sim_total$aggregate_cost),0), "\n")
cat("SD Aggregate Loss:   ", round(sd(sim_total$aggregate_cost),0), "\n")
cat("Median:              ", round(median(sim_total$aggregate_cost), 0), "\n")
cat("50%:                 ", round(quantile(sim_total$aggregate_cost, 0.50),0), "\n")
cat("90%:                 ", round(quantile(sim_total$aggregate_cost, 0.90),0), "\n")
cat("95%:                 ", round(quantile(sim_total$aggregate_cost, 0.95),0), "\n")
cat("99%:                 ", round(quantile(sim_total$aggregate_cost, 0.99),0), "\n")
cat("99.95%:               ", round(quantile(sim_total$aggregate_cost, 0.9995),0), "\n")
cat("TVaR 95%:            ", round(mean(sim_total$aggregate_cost[sim_total$aggregate_cost > quantile(sim_total$aggregate_cost, 0.95)]),0),"\n")
cat("TVaR 99%:            ", round(mean(sim_total$aggregate_cost[sim_total$aggregate_cost > quantile(sim_total$aggregate_cost, 0.99)]),0), "\n")
cat("TVaR 99.95%:          ", round(mean(sim_total$aggregate_cost[sim_total$aggregate_cost > quantile(sim_total$aggregate_cost, 0.9995)]),0), "\n")

# Net Revenue 
cat("\n--- Net Revenue ---\n")
cat("Mean Net Revenue:    ", round(mean(sim_total$net_revenue),0), "\n")
cat("SD Net Revenue:      ", round(sd(sim_total$net_revenue),0), "\n")
cat("Median:              ", round(median(sim_total$net_revenue),0), "\n")
cat("50%:                 ", round(quantile(sim_total$net_revenue, 0.50),0), "\n")
cat("90%:                 ", round(quantile(sim_total$net_revenue, 0.10),0), "\n")
cat("95%:                 ", round(quantile(sim_total$net_revenue, 0.05),0), "\n")
cat("99%:                 ", round(quantile(sim_total$net_revenue, 0.01),0), "\n")
cat("99.95%:               ", round(quantile(sim_total$net_revenue, 0.0005),0), "\n")

# Risk Metrics 
cat("\n--- Risk Metrics ---\n")
cat("Prob of Loss:        ", round(mean(sim_total$net_revenue < 0) * 100, 4), "%\n")
cat("Mean Loss Ratio:     ", round(mean(sim_total$loss_ratio) * 100, 2),     "%\n")
cat("Median Loss Ratio:   ", round(median(sim_total$loss_ratio) * 100, 2),   "%\n")

# High Value Segment 
cat("\n--- High Value Segment (Gold & Platinum) ---\n")
cat("Mean Aggregate Loss: ", round(mean(sim_high$aggregate_cost),0), "\n")
cat("Mean Net Revenue:    ", round(mean(sim_high$net_revenue),0), "\n")
cat("Prob of Loss:        ", round(mean(sim_high$net_revenue < 0) * 100, 4), "%\n")

# Standard Segment 
cat("\n--- Standard Segment ---\n")
cat("Mean Aggregate Loss: ", round(mean(sim_std$aggregate_cost), 0), "\n")
cat("Mean Net Revenue:    ", round(mean(sim_std$net_revenue),0), "\n")
cat("Prob of Loss:        ", round(mean(sim_std$net_revenue < 0) * 100, 4), "%\n")

# Sanity Check 
cat("\n--- Sanity Check ---\n")
cat("High + Std Mean Loss:  ", round(mean(sim_high$aggregate_cost) + mean(sim_std$aggregate_cost),0), "\n")
cat("Total Mean Loss:       ", round(mean(sim_total$aggregate_cost),0), "\n")
cat("Match:                 ", round(mean(sim_high$aggregate_cost) + mean(sim_std$aggregate_cost), 0) == round(mean(sim_total$aggregate_cost), 0),"\n")
```
# PLOT SIMULATION DISTRIBUTIONS
```{r}
par(mfrow = c(1, 3))

hist(sim_total$aggregate_cost,
     breaks = 50,
     main   = "Aggregate Cost Distribution",
     xlab   = "Aggregate Cost",
     col    = "steelblue")

hist(sim_total$revenue,
     breaks = 30,
     main   = "Revenue Distribution",
     xlab   = "Revenue",
     col    = "goldenrod")

hist(sim_total$net_revenue,
     breaks = 50,
     main   = "Net Revenue Distribution",
     xlab   = "Net Revenue",
     col    = "seagreen")
```

# LOADING SENSITIVITY ANALYSIS 
```{r}
set.seed(123)

loadings <- c(0.10, 0.20, 0.30, 0.40, 0.50)

loading_results <- do.call(rbind, lapply(loadings, function(l) {
  
  sim_high_l <- simulate_segment(
    freq_pred   = nb_high$freq_pred,
    sev_pred    = nb_high$sev_pred,
    theta       = theta,
    gamma_shape = gamma_shape,
    loading     = l,
    n_sim       = 10000
  )
  
  sim_std_l <- simulate_segment(
    freq_pred   = nb_std$freq_pred,
    sev_pred    = nb_std$sev_pred,
    theta       = theta,
    gamma_shape = gamma_shape,
    loading     = l,
    n_sim       = 10000
  )
  
  sim_total_l <- data.frame(
    aggregate_cost = sim_high_l$aggregate_cost + sim_std_l$aggregate_cost,
    revenue        = sim_high_l$revenue + sim_std_l$revenue,
    net_revenue    = sim_high_l$net_revenue + sim_std_l$net_revenue,
    loss_ratio     = (sim_high_l$aggregate_cost + sim_std_l$aggregate_cost) /
                     (sim_high_l$revenue + sim_std_l$revenue)
  )
  
  data.frame(
    Loading           = paste0(l * 100, "%"),
    Mean_Revenue      = round(mean(sim_total_l$revenue), 0),
    Mean_Cost         = round(mean(sim_total_l$aggregate_cost), 0),
    Mean_NetRevenue   = round(mean(sim_total_l$net_revenue), 0),
    Prob_Loss         = round(mean(sim_total_l$net_revenue < 0), 4),
    Mean_LossRatio    = round(mean(sim_total_l$loss_ratio), 4),
    VaR99_Cost        = round(quantile(sim_total_l$aggregate_cost, 0.99), 0),
    TVaR99_Cost       = round(mean(sim_total_l$aggregate_cost[
                          sim_total_l$aggregate_cost >
                          quantile(sim_total_l$aggregate_cost, 0.99)]), 0),
    VaR99_NetRevenue  = round(quantile(sim_total_l$net_revenue, 0.01), 0)
  )
}))

print(loading_results)
```

# Long Term Simulation
# Economic assumptions 
```{r}
avg_inflation <- 0.025   # 2.5%
avg_discount  <- 0.018   # 1.8%

cat("Inflation rate:", round(avg_inflation * 100, 2), "%\n")
cat("Discount rate:",  round(avg_discount  * 100, 2), "%\n")
```

# Fleet growth assumptions
```{r}
# solar growth assumptions 
annual_growth_helionis <- (1 + 0.25)^(1/10) - 1
annual_growth_bayesia  <- (1 + 0.25)^(1/10) - 1
annual_growth_oryn     <- (1 + 0.15)^(1/10) - 1

# solar sizes 
base_helionis <- 1160
base_bayesia  <- 1128
base_oryn     <- 774
```

# container proportion
```{r}
prop_helionis <- c(
  DeepSpace    = 58/1160,
  DockArc      = 116/1160,
  HardSeal     = 580/1160,
  LongHaul     = 232/1160,
  QuantumCrate = 174/1160
)

prop_bayesia <- c(
  DeepSpace    = 56/1128,
  DockArc      = 113/1128,
  HardSeal     = 564/1128,
  LongHaul     = 226/1128,
  QuantumCrate = 169/1128
)

prop_oryn <- c(
  DeepSpace    = 39/774,
  DockArc      = 77/774,
  HardSeal     = 387/774,
  LongHaul     = 155/774,
  QuantumCrate = 116/774
)
```

# Long term NB Generator
Assumption: The proportion of each container type remains constant as the fleet grows
```{r}
make_nb_data_cargo_lt <- function(df_freq, df_sev,
                                  n_helionis, n_bayesia, n_oryn,
                                  prop_helionis, prop_bayesia, prop_oryn) {
  
  total_n <- n_helionis + n_bayesia + n_oryn
  
  # Build container mixyup 
  container_helionis <- c(
    rep("DeepSpace Haulbox",       round(prop_helionis["DeepSpace"]    * n_helionis)),
    rep("DockArc Freight Case",    round(prop_helionis["DockArc"]      * n_helionis)),
    rep("HardSeal Transit Crate",  round(prop_helionis["HardSeal"]     * n_helionis)),
    rep("LongHaul Vault Canister", round(prop_helionis["LongHaul"]     * n_helionis)),
    rep("QuantumCrate Module",     round(prop_helionis["QuantumCrate"] * n_helionis))
  )
  
  container_bayesia <- c(
    rep("DeepSpace Haulbox",       round(prop_bayesia["DeepSpace"]    * n_bayesia)),
    rep("DockArc Freight Case",    round(prop_bayesia["DockArc"]      * n_bayesia)),
    rep("HardSeal Transit Crate",  round(prop_bayesia["HardSeal"]     * n_bayesia)),
    rep("LongHaul Vault Canister", round(prop_bayesia["LongHaul"]     * n_bayesia)),
    rep("QuantumCrate Module",     round(prop_bayesia["QuantumCrate"] * n_bayesia))
  )
  
  container_oryn <- c(
    rep("DeepSpace Haulbox",       round(prop_oryn["DeepSpace"]    * n_oryn)),
    rep("DockArc Freight Case",    round(prop_oryn["DockArc"]      * n_oryn)),
    rep("HardSeal Transit Crate",  round(prop_oryn["HardSeal"]     * n_oryn)),
    rep("LongHaul Vault Canister", round(prop_oryn["LongHaul"]     * n_oryn)),
    rep("QuantumCrate Module",     round(prop_oryn["QuantumCrate"] * n_oryn))
  )
  
  all_containers <- c(container_helionis, container_bayesia, container_oryn)
  
  # Trim for rounding mismatch 
  if (length(all_containers) > total_n) {
    all_containers <- all_containers[1:total_n]
  } else if (length(all_containers) < total_n) {
    all_containers <- c(all_containers,
                        rep("HardSeal Transit Crate", total_n - length(all_containers)))
  }
  
  # Row-based bootstrap from historical severity data
  sampled_rows <- df_sev[sample(seq_len(nrow(df_sev)), total_n, replace = TRUE), ]
  
  nb_data <- data.frame(
    cargo_value      = sampled_rows$cargo_value,
    pilot_experience = sampled_rows$pilot_experience,
    route_risk       = sampled_rows$route_risk,
    cargo_type       = sampled_rows$cargo_type,
    weight           = sampled_rows$weight,
    solar_radiation  = sampled_rows$solar_radiation,
    debris_density   = sampled_rows$debris_density,
    exposure         = 1
  )
  
  nb_data$container_type <- factor(
    all_containers,
    levels = levels(df_freq$container_type)
  )
  
  nb_data$cargo_type <- factor(nb_data$cargo_type,
                               levels = levels(df_freq$cargo_type))
  
  nb_data$route_risk <- factor(nb_data$route_risk,
                               levels = levels(df_freq$route_risk),
                               ordered = TRUE)
  
  nb_data$solar_system <- c(
    rep("Helionis Cluster", n_helionis),
    rep("Bayesia System",   n_bayesia),
    rep("Oryn Delta",       n_oryn)
  )
  
   # Cap weight by container type capacity 
  container_max_weight <- c(
    "DeepSpace Haulbox"       = 25000,
    "DockArc Freight Case"    = 50000,
    "HardSeal Transit Crate"  = 100000,
    "LongHaul Vault Canister" = 150000,
    "QuantumCrate Module"     = 250000
  )
  
  nb_data <- nb_data %>%
    mutate(
      weight = pmin(weight,
                    container_max_weight[as.character(container_type)])
    )
  
  nb_data
}
```

# Parameters 
```{r}
set.seed(123)

n_sim   <- 100000
n_years <- 10
loading <- 0.40

theta       <- model_cargo_freq3$theta
gamma_disp  <- summary(model_cargo_sev2)$dispersion
gamma_shape <- 1 / gamma_disp
```

# Pre-build all 10 years of new business data
```{r}
cat("Building new business data for all 10 years...\n")

yearly_nb_data <- list()

for (yr in 1:n_years) {
  
  cat("  Building year", yr, "data...\n")
  
  # Fleet size this year
  n_hel <- round(base_helionis * (1 + annual_growth_helionis)^yr)
  n_bay <- round(base_bayesia  * (1 + annual_growth_bayesia)^yr)
  n_ory <- round(base_oryn     * (1 + annual_growth_oryn)^yr)
  
  # Generate portfolio
  nb_data_yr <- make_nb_data_cargo_lt(
    df_freq       = df_cargo_freq,
    df_sev        = df_cargo_sev,
    n_helionis    = n_hel,
    n_bayesia     = n_bay,
    n_oryn        = n_ory,
    prop_helionis = prop_helionis,
    prop_bayesia  = prop_bayesia,
    prop_oryn     = prop_oryn
  )
  
  # Predictions
  nb_data_yr$freq_pred <- predict(model_cargo_freq3,
                                  newdata = nb_data_yr,
                                  type    = "response")
  nb_data_yr$sev_pred  <- predict(model_cargo_sev2,
                                  newdata = nb_data_yr,
                                  type    = "response")
  
  # Apply inflation to severity
  inflation_factor     <- (1 + avg_inflation)^yr
  nb_data_yr$sev_pred  <- nb_data_yr$sev_pred * inflation_factor
  
  # Discount factor
  discount_factor <- 1 / (1 + avg_discount)^yr
  
  # Pre-calculate annual earnings (fixed premium)
  annual_earnings <- sum(nb_data_yr$freq_pred * nb_data_yr$sev_pred) * (1 + loading)
  
  # Store
  yearly_nb_data[[yr]] <- list(
    nb_data          = nb_data_yr,
    n_exposures      = nrow(nb_data_yr),
    inflation_factor = inflation_factor,
    discount_factor  = discount_factor,
    annual_earnings  = annual_earnings
  )
}

cat("All 10 years data built! Now running joint simulation...\n")
```

# Joint 10-year simulation
```{r}
joint_sim <- data.frame(
  total_cost        = numeric(n_sim),
  total_revenue     = numeric(n_sim),
  total_netrev      = numeric(n_sim),
  pv_total_cost     = numeric(n_sim),
  pv_total_revenue  = numeric(n_sim),
  pv_total_netrev   = numeric(n_sim)
)

cat("Running", n_sim, "joint simulations...\n")

for (i in 1:n_sim) {
  
  if (i %% 10000 == 0) cat("  Simulation run", i, "of", n_sim, "...\n")
  
  run_total_cost       <- 0
  run_total_revenue    <- 0
  run_total_netrev     <- 0
  run_pv_total_cost    <- 0
  run_pv_total_revenue <- 0
  run_pv_total_netrev  <- 0
  
  for (yr in 1:n_years) {
    
    pred_freq_yr    <- yearly_nb_data[[yr]]$nb_data$freq_pred
    pred_sev_yr     <- yearly_nb_data[[yr]]$nb_data$sev_pred
    annual_earnings <- yearly_nb_data[[yr]]$annual_earnings
    discount_factor <- yearly_nb_data[[yr]]$discount_factor
    n_policies_yr   <- yearly_nb_data[[yr]]$n_exposures
    
    # Simulate claim counts — NB for cargo
    claim_counts <- rnbinom(n_policies_yr,
                            mu   = pred_freq_yr,
                            size = theta)
    
    yr_total_loss <- 0
    for (j in 1:n_policies_yr) {
      if (claim_counts[j] > 0) {
        gamma_rate    <- gamma_shape / pred_sev_yr[j]
        claims        <- rgamma(claim_counts[j],
                                shape = gamma_shape,
                                rate  = gamma_rate)
        yr_total_loss <- yr_total_loss + sum(claims)
      }
    }
    
    yr_netrev <- annual_earnings - yr_total_loss
    
    # Accumulate nominal
    run_total_cost    <- run_total_cost    + yr_total_loss
    run_total_revenue <- run_total_revenue + annual_earnings
    run_total_netrev  <- run_total_netrev  + yr_netrev
    
    # Accumulate PV
    run_pv_total_cost    <- run_pv_total_cost    + yr_total_loss   * discount_factor
    run_pv_total_revenue <- run_pv_total_revenue + annual_earnings * discount_factor
    run_pv_total_netrev  <- run_pv_total_netrev  + yr_netrev       * discount_factor
  }
  
  joint_sim$total_cost[i]       <- run_total_cost
  joint_sim$total_revenue[i]    <- run_total_revenue
  joint_sim$total_netrev[i]     <- run_total_netrev
  joint_sim$pv_total_cost[i]    <- run_pv_total_cost
  joint_sim$pv_total_revenue[i] <- run_pv_total_revenue
  joint_sim$pv_total_netrev[i]  <- run_pv_total_netrev
}

cat("Joint simulation complete!\n")
```

# Results output
```{r}
cat("\n=== LONG TERM CARGO — JOINT SIMULATION RESULTS ===\n")
cat("Number of simulations:", n_sim, "\n")
cat("Number of years:      ", n_years, "\n")
cat("Loading:               40%\n")

# Average and total fleet size
avg_fleet  <- round(mean(sapply(yearly_nb_data, function(x) x$n_exposures)), 0)
total_fleet <- round(sum(sapply(yearly_nb_data, function(x) x$n_exposures)), 0)

cat("Average Annual Fleet Size:", avg_fleet, "\n")
cat("Total Fleet Size (10 yr): ", total_fleet, "\n")

# NOMINAL
cat("\n=== NOMINAL (Future Value) ===\n")

cat("\n--- Aggregate Loss ---\n")
cat("Mean:     ", round(mean(joint_sim$total_cost),              0), "\n")
cat("Median:   ", round(median(joint_sim$total_cost),            0), "\n")
cat("StdDev:   ", round(sd(joint_sim$total_cost),                0), "\n")
cat("50%:      ", round(quantile(joint_sim$total_cost, 0.50),    0), "\n")
cat("90%:      ", round(quantile(joint_sim$total_cost, 0.90),    0), "\n")
cat("95%:      ", round(quantile(joint_sim$total_cost, 0.95),    0), "\n")
cat("99%:      ", round(quantile(joint_sim$total_cost, 0.99),    0), "\n")
cat("99.95%:    ", round(quantile(joint_sim$total_cost, 0.9995),   0), "\n")
cat("TVaR 99%: ", round(mean(joint_sim$total_cost[
                  joint_sim$total_cost >
                  quantile(joint_sim$total_cost, 0.99)]),         0), "\n")
cat("TVaR 99.95%:", round(mean(joint_sim$total_cost[
                  joint_sim$total_cost >
                  quantile(joint_sim$total_cost, 0.9995)]),        0), "\n")

cat("\n--- Revenue (Premium) ---\n")
cat("Mean:     ", round(mean(joint_sim$total_revenue),           0), "\n")
cat("Median:   ", round(median(joint_sim$total_revenue),         0), "\n")

cat("\n--- Net Revenue ---\n")
cat("Mean:     ", round(mean(joint_sim$total_netrev),              0), "\n")
cat("Median:   ", round(median(joint_sim$total_netrev),            0), "\n")
cat("StdDev:   ", round(sd(joint_sim$total_netrev),                0), "\n")
cat("50%:      ", round(quantile(joint_sim$total_netrev, 0.50),    0), "\n")
cat("90%:      ", round(quantile(joint_sim$total_netrev, 0.10),    0), "\n")
cat("95%:      ", round(quantile(joint_sim$total_netrev, 0.05),    0), "\n")
cat("99%:      ", round(quantile(joint_sim$total_netrev, 0.01),    0), "\n")
cat("99.95%:    ", round(quantile(joint_sim$total_netrev, 0.0005),   0), "\n")
cat("Prob of Loss: ", round(mean(joint_sim$total_netrev < 0), 4),  "\n")

# PRESENT VALUE 
cat("\n=== PRESENT VALUE (Today's Dollars) ===\n")

cat("\n--- PV Aggregate Loss ---\n")
cat("Mean:     ", round(mean(joint_sim$pv_total_cost),              0), "\n")
cat("Median:   ", round(median(joint_sim$pv_total_cost),            0), "\n")
cat("StdDev:   ", round(sd(joint_sim$pv_total_cost),                0), "\n")
cat("50%:      ", round(quantile(joint_sim$pv_total_cost, 0.50),    0), "\n")
cat("90%:      ", round(quantile(joint_sim$pv_total_cost, 0.90),    0), "\n")
cat("95%:      ", round(quantile(joint_sim$pv_total_cost, 0.95),    0), "\n")
cat("99%:      ", round(quantile(joint_sim$pv_total_cost, 0.99),    0), "\n")
cat("99.95%:    ", round(quantile(joint_sim$pv_total_cost, 0.9995),   0), "\n")
cat("TVaR 99%: ", round(mean(joint_sim$pv_total_cost[
                  joint_sim$pv_total_cost >
                  quantile(joint_sim$pv_total_cost, 0.99)]),         0), "\n")
cat("TVaR 99.95%:", round(mean(joint_sim$pv_total_cost[
                  joint_sim$pv_total_cost >
                  quantile(joint_sim$pv_total_cost, 0.9995)]),        0), "\n")

cat("\n--- PV Revenue (Premium) ---\n")
cat("Mean:     ", round(mean(joint_sim$pv_total_revenue),           0), "\n")
cat("Median:   ", round(median(joint_sim$pv_total_revenue),         0), "\n")

cat("\n--- PV Net Revenue ---\n")
cat("Mean:     ", round(mean(joint_sim$pv_total_netrev),              0), "\n")
cat("Median:   ", round(median(joint_sim$pv_total_netrev),            0), "\n")
cat("StdDev:   ", round(sd(joint_sim$pv_total_netrev),                0), "\n")
cat("50%:      ", round(quantile(joint_sim$pv_total_netrev, 0.50),    0), "\n")
cat("90%:      ", round(quantile(joint_sim$pv_total_netrev, 0.10),    0), "\n")
cat("95%:      ", round(quantile(joint_sim$pv_total_netrev, 0.05),    0), "\n")
cat("99%:      ", round(quantile(joint_sim$pv_total_netrev, 0.01),    0), "\n")
cat("99.95%:    ", round(quantile(joint_sim$pv_total_netrev, 0.0005),   0), "\n")
cat("Prob of Loss: ", round(mean(joint_sim$pv_total_netrev < 0), 4),  "\n")

# FOR TABLE USE
cat("\n=== FOR TABLE ===\n")

cat("\n--- NOMINAL ---\n")
cat("Risk Count:  ", round(mean(sapply(yearly_nb_data,
                    function(x) x$n_exposures)), 0), "\n")
cat("Loading:      40%\n")
cat("Premium:     ", round(mean(joint_sim$total_revenue),         0), "\n")
cat("\nAggregate Loss:\n")
cat("  Mean:      ", round(mean(joint_sim$total_cost),            0), "\n")
cat("  50%:       ", round(quantile(joint_sim$total_cost, 0.50),  0), "\n")
cat("  90%:       ", round(quantile(joint_sim$total_cost, 0.90),  0), "\n")
cat("  95%:       ", round(quantile(joint_sim$total_cost, 0.95),  0), "\n")
cat("  99%:       ", round(quantile(joint_sim$total_cost, 0.99),  0), "\n")
cat("  99.95%:     ", round(quantile(joint_sim$total_cost, 0.9995), 0), "\n")
cat("\nNet Revenue:\n")
cat("  Mean:      ", round(mean(joint_sim$total_netrev),            0), "\n")
cat("  50%:       ", round(quantile(joint_sim$total_netrev, 0.50),  0), "\n")
cat("  90%:       ", round(quantile(joint_sim$total_netrev, 0.10),  0), "\n")
cat("  95%:       ", round(quantile(joint_sim$total_netrev, 0.05),  0), "\n")
cat("  99%:       ", round(quantile(joint_sim$total_netrev, 0.01),  0), "\n")
cat("  99.95%:     ", round(quantile(joint_sim$total_netrev, 0.0005), 0), "\n")
cat("Prob of Loss:", round(mean(joint_sim$total_netrev < 0), 4),    "\n")

cat("\n--- PRESENT VALUE ---\n")
cat("PV Premium:  ", round(mean(joint_sim$pv_total_revenue),         0), "\n")
cat("\nPV Aggregate Loss:\n")
cat("  Mean:      ", round(mean(joint_sim$pv_total_cost),            0), "\n")
cat("  50%:       ", round(quantile(joint_sim$pv_total_cost, 0.50),  0), "\n")
cat("  90%:       ", round(quantile(joint_sim$pv_total_cost, 0.90),  0), "\n")
cat("  95%:       ", round(quantile(joint_sim$pv_total_cost, 0.95),  0), "\n")
cat("  99%:       ", round(quantile(joint_sim$pv_total_cost, 0.99),  0), "\n")
cat("  99.95%:     ", round(quantile(joint_sim$pv_total_cost, 0.9995), 0), "\n")
cat("\nPV Net Revenue:\n")
cat("  Mean:      ", round(mean(joint_sim$pv_total_netrev),            0), "\n")
cat("  50%:       ", round(quantile(joint_sim$pv_total_netrev, 0.50),  0), "\n")
cat("  90%:       ", round(quantile(joint_sim$pv_total_netrev, 0.10),  0), "\n")
cat("  95%:       ", round(quantile(joint_sim$pv_total_netrev, 0.05),  0), "\n")
cat("  99%:       ", round(quantile(joint_sim$pv_total_netrev, 0.01),  0), "\n")
cat("  99.95%:     ", round(quantile(joint_sim$pv_total_netrev, 0.0005), 0), "\n")
cat("PV Prob of Loss:", round(mean(joint_sim$pv_total_netrev < 0), 4), "\n")
```

# Results output visualisation
```{r}
# Plots 
par(mfrow = c(2, 2))

# Plot 1: Nominal aggregate loss distribution
hist(joint_sim$total_cost / 1e9,
     breaks = 50,
     main   = "10-Year Nominal Aggregate Loss",
     xlab   = "Total Loss (Billions)",
     col    = "steelblue",
     border = "white")
abline(v = mean(joint_sim$total_cost) / 1e9,
       col = "blue", lwd = 2, lty = 2)
abline(v = quantile(joint_sim$total_cost, 0.99) / 1e9,
       col = "red", lwd = 2, lty = 2)
abline(v = quantile(joint_sim$total_cost, 0.995) / 1e9,
       col = "darkred", lwd = 2, lty = 2)
legend("topright",
       legend = c("Mean", "VaR 99%", "VaR 99.5%"),
       col    = c("blue", "red", "darkred"),
       lty = 2, lwd = 2, cex = 0.8)

# Plot 2: Nominal net revenue distribution
hist(joint_sim$total_netrev / 1e9,
     breaks = 50,
     main   = "10-Year Nominal Net Revenue",
     xlab   = "Net Revenue (Billions)",
     col    = "seagreen",
     border = "white")
abline(v = mean(joint_sim$total_netrev) / 1e9,
       col = "darkgreen", lwd = 2, lty = 2)
abline(v = 0,
       col = "black", lwd = 2, lty = 1)
legend("topright",
       legend = c("Mean", "Break Even"),
       col    = c("darkgreen", "black"),
       lty    = c(2, 1), lwd = 2, cex = 0.8)

# Plot 3: PV aggregate loss distribution
hist(joint_sim$pv_total_cost / 1e9,
     breaks = 50,
     main   = "10-Year PV Aggregate Loss",
     xlab   = "PV Total Loss (Billions)",
     col    = "tomato",
     border = "white")
abline(v = mean(joint_sim$pv_total_cost) / 1e9,
       col = "red", lwd = 2, lty = 2)
abline(v = quantile(joint_sim$pv_total_cost, 0.99) / 1e9,
       col = "darkred", lwd = 2, lty = 2)
legend("topright",
       legend = c("Mean", "VaR 99%"),
       col    = c("red", "darkred"),
       lty = 2, lwd = 2, cex = 0.8)

# Plot 4: PV net revenue distribution
hist(joint_sim$pv_total_netrev / 1e9,
     breaks = 50,
     main   = "10-Year PV Net Revenue",
     xlab   = "PV Net Revenue (Billions)",
     col    = "goldenrod",
     border = "white")
abline(v = mean(joint_sim$pv_total_netrev) / 1e9,
       col = "darkgoldenrod", lwd = 2, lty = 2)
abline(v = 0,
       col = "black", lwd = 2, lty = 1)
legend("topright",
       legend = c("Mean", "Break Even"),
       col    = c("darkgoldenrod", "black"),
       lty    = c(2, 1), lwd = 2, cex = 0.8)
```

# Stress Testing for short term
# PARAMETERS
```{r}
set.seed(123)

# Parameters 
loading <- 0.40
n_sim   <- 100000

# Segment flag 
nb_data_cargo$segment <- ifelse(
  nb_data_cargo$cargo_type %in% c("gold", "platinum"),
  "high_value",
  "standard"
)

nb_high <- subset(nb_data_cargo, segment == "high_value")
nb_std  <- subset(nb_data_cargo, segment == "standard")

# Model parameters 
theta       <- model_cargo_freq3$theta
gamma_disp  <- summary(model_cargo_sev2)$dispersion
gamma_shape <- 1 / gamma_disp
```

# BASE PREM SETUP
```{r}
base_freq_high <- nb_high$freq_pred
base_sev_high  <- nb_high$sev_pred

base_freq_std  <- nb_std$freq_pred
base_sev_std   <- nb_std$sev_pred

base_premium_high <- sum(base_freq_high * base_sev_high) * (1 + loading)
base_premium_std  <- sum(base_freq_std  * base_sev_std)  * (1 + loading)
base_premium_total <- base_premium_high + base_premium_std
```

# SIMULATION FUNCTION WITH FIXED PREMIUM
```{r}
simulate_segment_fixed_premium <- function(freq_pred, sev_pred, theta, gamma_shape,
                                           fixed_premium, n_sim) {
  
  n_units <- length(freq_pred)
  
  results <- data.frame(
    aggregate_cost = numeric(n_sim),
    revenue        = rep(fixed_premium, n_sim),
    net_revenue    = numeric(n_sim),
    loss_ratio     = numeric(n_sim)
  )
  
  for (i in 1:n_sim) {
    
    claim_counts <- rnbinom(n_units, mu = freq_pred, size = theta)
    total_loss <- 0
    
    for (j in 1:n_units) {
      if (claim_counts[j] > 0) {
        gamma_rate <- gamma_shape / sev_pred[j]
        claims <- rgamma(claim_counts[j], shape = gamma_shape, rate = gamma_rate)
        total_loss <- total_loss + sum(claims)
      }
    }
    
    results$aggregate_cost[i] <- total_loss
    results$net_revenue[i]    <- fixed_premium - total_loss
    results$loss_ratio[i]     <- total_loss / fixed_premium
  }
  
  return(results)
}
```

# STRESS-TEST WRAPPER
```{r}
run_cargo_stress_1yr <- function(freq_stress = 1.00,
                                 sev_stress  = 1.00,
                                 scenario_name = "Base") {
  
  # Apply stress to predicted outputs
  stressed_freq_high <- base_freq_high * freq_stress
  stressed_sev_high  <- base_sev_high  * sev_stress
  
  stressed_freq_std  <- base_freq_std * freq_stress
  stressed_sev_std   <- base_sev_std  * sev_stress
  
  # Simulate each segment with FIXED base premium
  sim_high <- simulate_segment_fixed_premium(
    freq_pred     = stressed_freq_high,
    sev_pred      = stressed_sev_high,
    theta         = theta,
    gamma_shape   = gamma_shape,
    fixed_premium = base_premium_high,
    n_sim         = n_sim
  )
  
  sim_std <- simulate_segment_fixed_premium(
    freq_pred     = stressed_freq_std,
    sev_pred      = stressed_sev_std,
    theta         = theta,
    gamma_shape   = gamma_shape,
    fixed_premium = base_premium_std,
    n_sim         = n_sim
  )
  
  # Combine segment results into total portfolio
  sim_total <- data.frame(
    aggregate_cost = sim_high$aggregate_cost + sim_std$aggregate_cost,
    revenue        = sim_high$revenue + sim_std$revenue,
    net_revenue    = sim_high$net_revenue + sim_std$net_revenue,
    loss_ratio     = (sim_high$aggregate_cost + sim_std$aggregate_cost) /
                     (sim_high$revenue + sim_std$revenue)
  )
  
  # Tail metrics — cost side
  var50_loss  <- as.numeric(quantile(sim_total$aggregate_cost, 0.50))
  var90_loss  <- as.numeric(quantile(sim_total$aggregate_cost, 0.90))
  var95_loss  <- as.numeric(quantile(sim_total$aggregate_cost, 0.95))
  var99_loss  <- as.numeric(quantile(sim_total$aggregate_cost, 0.99))
  var995_loss <- as.numeric(quantile(sim_total$aggregate_cost, 0.9995))
  
  tvar95_loss  <- mean(sim_total$aggregate_cost[sim_total$aggregate_cost > var95_loss])
  tvar99_loss  <- mean(sim_total$aggregate_cost[sim_total$aggregate_cost > var99_loss])
  tvar995_loss <- mean(sim_total$aggregate_cost[sim_total$aggregate_cost > var995_loss])
  
  # Net revenue percentiles — left tail (worst outcomes)
  profit_50  <- as.numeric(quantile(sim_total$net_revenue, 0.50))
  profit_90  <- as.numeric(quantile(sim_total$net_revenue, 0.10))
  profit_95  <- as.numeric(quantile(sim_total$net_revenue, 0.05))
  profit_99  <- as.numeric(quantile(sim_total$net_revenue, 0.01))
  profit_995 <- as.numeric(quantile(sim_total$net_revenue, 0.0005))
  
  data.frame(
    Scenario              = scenario_name,
    N_Exposures           = nrow(nb_data_cargo),
    Fixed_Premium_Revenue = base_premium_total,
    Mean_Aggregate_Loss   = mean(sim_total$aggregate_cost),
    SD_Aggregate_Loss     = sd(sim_total$aggregate_cost),
    Mean_Net_Revenue      = mean(sim_total$net_revenue),
    SD_Net_Revenue        = sd(sim_total$net_revenue),
    Prob_Loss             = mean(sim_total$net_revenue < 0),
    Mean_Loss_Ratio       = mean(sim_total$loss_ratio),
    VaR_50_Loss           = var50_loss,
    VaR_90_Loss           = var90_loss,
    VaR_95_Loss           = var95_loss,
    VaR_99_Loss           = var99_loss,
    VaR_99_5_Loss         = var995_loss,
    TVaR_95_Loss          = tvar95_loss,
    TVaR_99_Loss          = tvar99_loss,
    TVaR_99_5_Loss        = tvar995_loss,
    Profit_50             = profit_50,
    Profit_90             = profit_90,
    Profit_95             = profit_95,
    Profit_99             = profit_99,
    Profit_99_5           = profit_995
  )
}
```

#  RUN STRESS SCENARIOS
```{r}
cargo_st_base <- run_cargo_stress_1yr(
  freq_stress   = 1.00,
  sev_stress    = 1.00,
  scenario_name = "Base"
)

cargo_st_freq <- run_cargo_stress_1yr(
  freq_stress   = 1.20,
  sev_stress    = 1.00,
  scenario_name = "Frequency +20%"
)

cargo_st_sev <- run_cargo_stress_1yr(
  freq_stress   = 1.00,
  sev_stress    = 1.20,
  scenario_name = "Severity +20%"
)

cargo_st_combined <- run_cargo_stress_1yr(
  freq_stress   = 1.20,
  sev_stress    = 1.20,
  scenario_name = "Freq +20% & Sev +20%"
)
```

# OUTPUT RESULTS
```{r}
cargo_stress_1yr_summary <- dplyr::bind_rows(
  cargo_st_base,
  cargo_st_freq,
  cargo_st_sev,
  cargo_st_combined
)
```

# DISPLAY RESULT
```{r}

cargo_stress_1yr_summary_print <- cargo_stress_1yr_summary

# Round all numeric columns first
cargo_stress_1yr_summary_print[, sapply(cargo_stress_1yr_summary_print, is.numeric)] <-
  round(cargo_stress_1yr_summary_print[, sapply(cargo_stress_1yr_summary_print, is.numeric)], 0)

# Then override ratios/probabilities with 4 decimals
cargo_stress_1yr_summary_print$Mean_Loss_Ratio <- round(cargo_stress_1yr_summary$Mean_Loss_Ratio, 4)
cargo_stress_1yr_summary_print$Prob_Loss       <- round(cargo_stress_1yr_summary$Prob_Loss, 4)

cat("\n=== STRESS TEST RESULTS - CARGO SHORT TERM (1 YEAR) ===\n")
cat("Base Premium Revenue:", round(base_premium_total, 0), "\n")
cat("Number of Exposures:", nrow(nb_data_cargo), "\n")
cat("Loading:", paste0(loading * 100, "%"), "\n\n")

print(cargo_stress_1yr_summary_print)
```

# DISPLAY COST and REVENUE
```{r}
# Table 1: Cost  
cost_table <- cargo_stress_1yr_summary %>%
  dplyr::select(Scenario, N_Exposures, Fixed_Premium_Revenue,
                Mean_Aggregate_Loss, VaR_50_Loss, VaR_90_Loss,
                VaR_95_Loss, VaR_99_Loss, VaR_99_5_Loss,
                TVaR_95_Loss, TVaR_99_Loss, TVaR_99_5_Loss)

cat("\n=== COST METRICS ===\n")
print(cost_table)

# Table 2: Net revenue 
revenue_table <- cargo_stress_1yr_summary %>%
  dplyr::select(Scenario, Fixed_Premium_Revenue,
                Mean_Net_Revenue, Profit_50, Profit_90,
                Profit_95, Profit_99, Profit_99_5,
                Prob_Loss, Mean_Loss_Ratio)

cat("\n=== NET REVENUE METRICS ===\n")
print(revenue_table)
```

# EXPORT AGGREGARE LOSS (SHORT TERM) FOR SCENARIO TESTING 
```{r}
#  Full results — all columns
saveRDS(
  sim_total,
  file = paste0(bronze_path, "cargo_st_base_full.rds")
)
cat("Full results saved!\n")
cat("Rows:", nrow(sim_total), "\n")
cat("Cols:", ncol(sim_total), "\n")

# Aaron's version — aggregate loss only 
saveRDS(
  sim_total$aggregate_cost,
  file = paste0(bronze_path, "cargo_st_base_agg_loss.rds")
)
cat("Aggregate loss only saved!\n")
cat("Length:", length(sim_total$aggregate_cost), "\n")
```


