---
title: "GLM_EF_claim"
output: html_document
date: "2026-03-17"
---

```{r setup, include=FALSE}

# install.packages("lme4")
# install.packages("glmmTMB")
# install.packages("kableExtra")

library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
library(lme4)
library(glmmTMB)
library(kableExtra)
library(MASS)

bronze_path <- paste0(getwd(), '/Bronze/')

ef_freq_model_clean <- read_csv(paste0(bronze_path, "ef_claim_freq_update.csv"))
ef_sev_model_clean <- read_csv(paste0(bronze_path, "ef_claim_sev_update.csv"))
```

# CHECK DATA
```{r}
######## Range Checks — ef_freq_model_clean ########
cat("=== FREQ TABLE RANGE CHECKS ===\n")
# 1. claim_count (valid: 0-3)
cat("claim_count:", min(ef_freq_model_clean$claim_count), "to", max(ef_freq_model_clean$claim_count), "\n")
# 2. equipment_age (valid: 0-10)
cat("equipment_age:", min(ef_freq_model_clean$equipment_age), "to", max(ef_freq_model_clean$equipment_age), "\n")
# 3. maintenance_int (valid: 100-5000)
cat("maintenance_int:", min(ef_freq_model_clean$maintenance_int), "to", max(ef_freq_model_clean$maintenance_int), "\n")
# 4. usage_int (valid: 0-24)
cat("usage_int:", min(ef_freq_model_clean$usage_int), "to", max(ef_freq_model_clean$usage_int), "\n")
# 5. exposure (valid: 0-1)
cat("exposure:", min(ef_freq_model_clean$exposure), "to", max(ef_freq_model_clean$exposure), "\n")
# 6. solar_system (categorical)
cat("solar_system unique values:", sort(unique(ef_freq_model_clean$solar_system)), "\n")
# 7. equipment_type (categorical)
cat("equipment_type unique values:", sort(unique(ef_freq_model_clean$equipment_type)), "\n")

######## Range Checks — ef_sev_model_clean ########
cat("\n=== SEV TABLE RANGE CHECKS ===\n")
# 1. claim_amount (valid: ~11K-790K)
cat("claim_amount:", min(ef_sev_model_clean$claim_amount), "to", max(ef_sev_model_clean$claim_amount), "\n")
# 2. equipment_age (valid: 0-10)
cat("equipment_age:", min(ef_sev_model_clean$equipment_age), "to", max(ef_sev_model_clean$equipment_age), "\n")
# 3. maintenance_int (valid: 100-5000)
cat("maintenance_int:", min(ef_sev_model_clean$maintenance_int), "to", max(ef_sev_model_clean$maintenance_int), "\n")
# 4. usage_int (valid: 0-24)
cat("usage_int:", min(ef_sev_model_clean$usage_int), "to", max(ef_sev_model_clean$usage_int), "\n")
# 5. exposure (valid: 0-1)
cat("exposure:", min(ef_sev_model_clean$exposure), "to", max(ef_sev_model_clean$exposure), "\n")
# 6. solar_system (categorical)
cat("solar_system unique values:", sort(unique(ef_sev_model_clean$solar_system)), "\n")
# 7. equipment_type (categorical)
cat("equipment_type unique values:", sort(unique(ef_sev_model_clean$equipment_type)), "\n")

```

# DATA FORMATING
```{r}
# Casting DF — Freq table
df_ef_freq <- ef_freq_model_clean %>% mutate(
    equipment_age   = as.numeric(equipment_age),
    maintenance_int = as.numeric(maintenance_int),
    usage_int       = as.numeric(usage_int),
    exposure        = as.numeric(exposure),
    claim_count     = as.numeric(claim_count),
    equipment_type  = factor(equipment_type),
    solar_system    = factor(solar_system)
  ) %>%
  filter(!is.na(claim_count), exposure > 0)

# Casting DF — Sev table
df_ef_sev <- ef_sev_model_clean %>% mutate(
    equipment_age   = as.numeric(equipment_age),
    maintenance_int = as.numeric(maintenance_int),
    usage_int       = as.numeric(usage_int),
    exposure        = as.numeric(exposure),
    claim_amount    = as.numeric(claim_amount),
    equipment_type  = factor(equipment_type),
    solar_system    = factor(solar_system)
  ) %>%
  filter(!is.na(claim_amount), claim_amount > 0)

# Verify
nrow(df_ef_freq)
nrow(df_ef_sev)
glimpse(df_ef_freq)
```

# FREQ GLM FIT
```{r}
freq_model_ef <- glm(
  claim_count ~ equipment_type + equipment_age + solar_system + maintenance_int + usage_int,
          data   = df_ef_freq,
          family = poisson(link = "log"),
          offset = log(exposure))

summary(freq_model_ef)
```

# SEV GLM FIT
```{r}
# Severity Model — Gamma GLM 
sev_model_ef1 <- glm(claim_amount ~ equipment_type + solar_system + equipment_age + maintenance_int + usage_int,
  data   = df_ef_sev,
  family = Gamma(link = "log"))

summary(sev_model_ef1)
```
# TRIMMED SEV GLM FIT
```{r}
sev_model_ef <- glm(claim_amount ~ equipment_type + solar_system + equipment_age + usage_int,
  data   = df_ef_sev,
  family = Gamma(link = "log"))

summary(sev_model_ef)

```
Based on the description and the coefficient mapping for sev and fre: mapping assumption: Epsilon = Bayesia, Zeta = Oryn Delta (refer explaination)

# REMAP SOLAR SYSTEM
```{r}
df_ef_freq_remapped <- df_ef_freq %>% mutate(solar_system = 
    case_when(
      solar_system == "Epsilon"          ~ "Bayesia System",
      solar_system == "Zeta"             ~ "Oryn Delta",
      solar_system == "Helionis Cluster" ~ "Helionis Cluster",
      TRUE ~ solar_system),
      
    solar_system = factor(solar_system,
                          levels = c("Helionis Cluster", "Bayesia System", "Oryn Delta")))

df_ef_sev_remapped <- df_ef_sev %>% mutate(solar_system = 
    case_when(
      solar_system == "Epsilon"          ~ "Bayesia System",
      solar_system == "Zeta"             ~ "Oryn Delta",
      solar_system == "Helionis Cluster" ~ "Helionis Cluster",
      TRUE ~ solar_system),
      
    solar_system = factor(solar_system,
                          levels = c("Helionis Cluster", "Bayesia System", "Oryn Delta")))

# sense check
cat("FREQ remapped solar_system")
print(table(df_ef_freq_remapped$solar_system))
cat("ORIGINAL freq solar_system (unchanged)")
print(table(df_ef_freq$solar_system))
cat("SEV remapped solar_system")
print(table(df_ef_sev_remapped$solar_system))
cat("ORIGINAL SEV solar_system (unchanged)")
print(table(df_ef_sev$solar_system))
```

# REFIT FREQ AND SEV GLM AFTER SOLAR SYSTEM MAPPING
```{r}
freq_model_eff_final <- glm(claim_count ~ equipment_type + equipment_age + solar_system + maintenance_int + usage_int,
          data   = df_ef_freq_remapped,
          family = poisson(link = "log"),
          offset = log(exposure))
summary(freq_model_eff_final)

sev_model_ef_final <- glm(claim_amount ~ equipment_type + solar_system + equipment_age + usage_int,
  data   = df_ef_sev_remapped,
  family = Gamma(link = "log"))
summary(sev_model_ef_final)
```
# RESIDUAL PLOT 
```{r}
# Residual Diagnostics 
par(mfrow = c(2, 2))

plot(freq_model_eff_final, main = "Freq Mode")
plot(sev_model_ef_final, main = "Sev Model")
```
# EDA
```{r}
# Historical Data — Frequency Summary 
cat("=== HISTORICAL FREQUENCY BY EQUIPMENT TYPE ===\n")
df_ef_freq_remapped %>%
  group_by(equipment_type) %>%
  summarise(
    n_policies       = n(),
    total_claims     = sum(claim_count),
    total_exposure   = round(sum(exposure), 1),
    avg_claim_rate   = round(sum(claim_count) / sum(exposure), 4),
    avg_claim_count  = round(mean(claim_count), 4),
    .groups = "drop"
  ) %>%
  arrange(desc(avg_claim_rate)) %>%
  print(n = Inf, width = Inf)

# Historical Data — Severity Summary 
cat("\n=== HISTORICAL SEVERITY BY EQUIPMENT TYPE ===\n")
df_ef_sev_remapped %>%
  group_by(equipment_type) %>%
  summarise(
    n_claims   = n(),
    min_sev    = round(min(claim_amount),    0),
    max_sev    = round(max(claim_amount),    0),
    mean_sev   = round(mean(claim_amount),   0),
    median_sev = round(median(claim_amount), 0),
    .groups = "drop"
  ) %>%
  arrange(desc(mean_sev)) %>%
  print(n = Inf, width = Inf)



# Historical Data — By Solar System 
cat("\n=== HISTORICAL FREQUENCY BY SOLAR SYSTEM ===\n")
df_ef_freq_remapped %>%
  group_by(solar_system) %>%
  summarise(
    n_policies     = n(),
    total_claims   = sum(claim_count),
    total_exposure = round(sum(exposure), 1),
    avg_claim_rate = round(sum(claim_count) / sum(exposure), 4),
    .groups = "drop"
  ) %>%
  arrange(desc(avg_claim_rate)) %>%
  print(n = Inf, width = Inf)

cat("\n=== HISTORICAL SEVERITY BY SOLAR SYSTEM ===\n")
df_ef_sev_remapped %>%
  group_by(solar_system) %>%
  summarise(
    n_claims   = n(),
    mean_sev   = round(mean(claim_amount),   0),
    median_sev = round(median(claim_amount), 0),
    .groups = "drop"
  ) %>%
  arrange(desc(mean_sev)) %>%
  print(n = Inf, width = Inf)
```


# INPUT INVENTORY DATA
```{r}
# Equipment counts by solar and equipment 
inventory <- data.frame(
  solar_system = rep(c("Helionis Cluster", "Bayesia System", "Oryn Delta"), each = 6),
  equipment_type = rep(c("Quantum Bores", "Graviton Extractors", "Fexstram Carriers",
                         "ReglAggregators", "Flux Riders", "Ion Pulverizers"), 3),
  count = c(
    300, 240, 150, 300, 1500, 90,   # Helionis Cluster
    150, 120, 75, 150, 750, 45, # Bayesia System
    100, 80, 50, 100, 500, 30 # Oryn Delta
  )
)

# Usage % in operation by solar and equipment 
usage_pct <- data.frame(
  solar_system = rep(c("Helionis Cluster", "Bayesia System", "Oryn Delta"), each = 6),
  equipment_type = rep(c("Quantum Bores", "Graviton Extractors", "Fexstram Carriers",
                         "ReglAggregators", "Flux Riders", "Ion Pulverizers"), 3),
  usage_pct = c(
    0.95, 0.95, 0.90, 0.80, 0.80, 0.50,  # Helionis Cluster
    0.80, 0.80, 0.75, 0.75, 0.80, 0.60, # Bayesia System
    0.75, 0.75, 0.70, 0.70, 0.75, 0.50 # Oryn Delta
  )
)

# Maintenance schedule (hrs) by solar and equipment 
maintenance <- data.frame(
  solar_system = rep(c("Helionis Cluster", "Bayesia System", "Oryn Delta"), each = 6),
  equipment_type = rep(c("Quantum Bores", "Graviton Extractors", "Fexstram Carriers",
                         "ReglAggregators", "Flux Riders", "Ion Pulverizers"), 3),
  maintenance_hrs = c(
    750, 750, 375, 1500, 1500, 1000, # Helionis Cluster
    600, 600, 400, 1000, 1000, 750, # Bayesia System
    500, 500, 250, 300, 300, 500 # Oryn Delta
  )
)

# Average Risk Index by solar  and equipment 
risk_index <- data.frame(
  solar_system = rep(c("Helionis Cluster", "Bayesia System", "Oryn Delta"), each = 6),
  equipment_type = rep(c("Quantum Bores", "Graviton Extractors", "Fexstram Carriers",
                         "ReglAggregators", "Flux Riders", "Ion Pulverizers"), 3),
  risk_idx = c(
    0.69, 0.48, 0.67, 0.24, 0.24, 0.64, # Helionis Cluster
    0.77, 0.57, 0.71, 0.26, 0.21, 0.66, # Bayesia System
    0.93, 0.73, 0.78, 0.35, 0.20, 0.75 # Oryn Delta
  )
)

# Service years: solar x equip x age 
service_years <- data.frame(
  solar_system = rep(c("Helionis Cluster", "Bayesia System", "Oryn Delta"), each = 30), # each solar have 30times because 6 equip x 5 year bands 
  equipment_type = rep(rep(c("Quantum Bores", "Graviton Extractors", "Fexstram Carriers",
                              "ReglAggregators", "Flux Riders", "Ion Pulverizers"), each = 5), 3), # each equip have 5 rows for year bands and for each solar
  age_mid = rep(c(2.5, 7, 12, 17, 25), 18), # 6 quip x 3 solar
  n_equipment = c(
    # Helionis Cluster
    30, 45, 180, 30, 15,   # Quantum Bores
    24, 36, 144, 24, 12,   # Graviton Extractors
    15, 23,  89, 15,  8,   # Fexstram Carriers
    30, 45, 180, 30, 15,   # ReglAggregators
    150,225, 900,150, 75,  # Flux Riders
     9, 14,  53,  9,  5,   # Ion Pulverizers
    # Bayesia System
    45, 38, 59,  8,  0,    # Quantum Bores
    36, 30, 48,  6,  0,    # Graviton Extractors
    23, 19, 29,  4,  0,    # Fexstram Carriers
    45, 38, 59,  8,  0,    # ReglAggregators
    225,187,300, 38,  0,   # Flux Riders
    14, 11, 18,  2,  0,    # Ion Pulverizers
    # Oryn Delta
    75, 15, 10,  0,  0,    # Quantum Bores
    60, 12,  8,  0,  0,    # Graviton Extractors
    37,  8,  5,  0,  0,    # Fexstram Carriers
    75, 15, 10,  0,  0,    # ReglAggregators
    375, 75, 50,  0,  0,   # Flux Riders
    22,  5,  3,  0,  0     # Ion Pulverizers
  )
)
```
 Oryn Data operation only recently establish, so make sense has mostly 0s in the older bands 
 Helionis olderst system. running logest equipment across all bands, most matured at 10-14 band. 

# BUILD NB DATAFRAME
```{r}
set.seed(42)

new_business_ef <- service_years %>%
  filter(n_equipment > 0) %>%
  left_join(usage_pct,   by = c("solar_system", "equipment_type")) %>%
  left_join(maintenance, by = c("solar_system", "equipment_type")) %>%
  left_join(risk_index,  by = c("solar_system", "equipment_type")) %>%
  uncount(n_equipment) %>% # expand the rows 
  mutate(
    usage_int       = usage_pct * 24,      # Convert usage % to hours per day (0-24 scale)
    maintenance_int = maintenance_hrs,     # Maintenance interval directly from inventory
    equipment_age   = case_when(           # Random sampling within age bands for realistic variation
      age_mid == 2.5 ~ runif(n(), 0,  5),
      age_mid == 7   ~ runif(n(), 5,  9),
      age_mid == 12  ~ runif(n(), 10, 14),
      age_mid == 17  ~ runif(n(), 15, 19),
      age_mid == 25  ~ runif(n(), 20, 30),
      TRUE           ~ age_mid),
  
    exposure        = 1,   # Full year 
    
    # Standardise equipment_type names to match model factor levels
    equipment_type  = case_when(
      equipment_type == "Quantum Bores"       ~ "Quantum Bore",
      equipment_type == "Graviton Extractors" ~ "Graviton Extractor",
      equipment_type == "Fexstram Carriers"   ~ "FexStram Carrier",
      equipment_type == "ReglAggregators"     ~ "ReglAggregators",
      equipment_type == "Flux Riders"         ~ "Flux Rider",
      equipment_type == "Ion Pulverizers"     ~ "Ion Pulverizer",
      TRUE ~ equipment_type),
    
    equipment_type = factor(equipment_type,
                            levels = levels(df_ef_freq_remapped$equipment_type))) %>%
  
  dplyr::select(solar_system, equipment_type, equipment_age,
                maintenance_int, usage_int, exposure, risk_idx)

# check final NB Table 
cat("New business rows:", nrow(new_business_ef))

# check equipment_age
cat("Equipment age spread")
new_business_ef %>%
  group_by(solar_system) %>%
  summarise(
    min_age    = round(min(equipment_age),    1),
    max_age    = round(max(equipment_age),    1),
    mean_age   = round(mean(equipment_age),   1),
    median_age = round(median(equipment_age), 1),
    .groups = "drop"
  ) %>%
  print()


print(table(new_business_ef$solar_system))
print(table(new_business_ef$equipment_type))
cat("usage_int:", min(new_business_ef$usage_int), "to", max(new_business_ef$usage_int), "\n")
cat("maintenance_int:", min(new_business_ef$maintenance_int), "to", max(new_business_ef$maintenance_int), "\n")
cat("risk_idx:", min(new_business_ef$risk_idx), "to", max(new_business_ef$risk_idx), "\n")

# verify risk index map correctly
new_business_ef %>%
  group_by(solar_system, equipment_type) %>%
  summarise(risk_idx = first(risk_idx), .groups = "drop") %>%
  arrange(solar_system, equipment_type) %>%
  print(n = Inf)
```

Final freq model :freq_model_eff_final
Final sev model :sev_model_ef_final

# NB PREDICTION
```{r}
# Predict base freq 
new_business_ef <- new_business_ef %>% mutate(pred_freq = predict(freq_model_eff_final,newdata  = new_business_ef,type  = "response"),)

# Predict base sev 
new_business_ef <- new_business_ef %>%mutate( pred_sev = predict(sev_model_ef_final, newdata = new_business_ef, type    = "response"))

# Expected pure premium per equipment unit
# Pure premium = expected frequency × expected severity
new_business_ef <- new_business_ef %>% mutate( pure_premium = pred_freq * pred_sev)

# NB prediction values - split by equip x solar 
# Summary by solar system 
new_business_ef %>%
  group_by(solar_system) %>%
  summarise(
    n_equipment       = n(),
    avg_pred_freq     = round(mean(pred_freq),     4),
    avg_pred_sev      = round(mean(pred_sev),      0),
    avg_pure_premium  = round(mean(pure_premium),  0),
    total_exp_losses  = round(sum(pure_premium),   0),
    .groups = "drop"
  ) %>%
  print()

# Summary by equipment type 
new_business_ef %>%
  group_by(equipment_type) %>%
  summarise(
    n_equipment      = n(),
    avg_pred_freq    = round(mean(pred_freq),    4),
    avg_pred_sev     = round(mean(pred_sev),     0),
    avg_pure_premium = round(mean(pure_premium), 0),
    total_exp_losses = round(sum(pure_premium),  0),
    .groups = "drop"
  ) %>%
  arrange(desc(total_exp_losses)) %>%
  print()

# Summary by solar system × equipment type
new_business_ef %>%
  group_by(solar_system, equipment_type) %>%
  summarise(
    n_equipment      = n(),
    avg_pred_freq    = round(mean(pred_freq),    4),
    avg_pred_sev     = round(mean(pred_sev),     0),
    total_exp_losses = round(sum(pure_premium),  0),
    .groups = "drop"
  ) %>%
  arrange(solar_system, desc(total_exp_losses)) %>%
  print(n = Inf)

# Overall total expected losses
cat("Total expected losses (1 year):", round(sum(new_business_ef$pure_premium), 0), "\n")
```
Intepretaion: 
NB total expected loss: 100600k
By freq : Helionis --> Bayesia --> Oryn, Sev: Oryn --> Bayesia --> Helionis
By Equipment: Flux --> Quantum --> Graviton --> reG --> Fex --> Ion
Quantum Bores represent the highest risk per unit despite lower frequency — their average claim of 142,025 is more than double FexStram Carriers. Flux Riders drive total losses primarily through fleet size.
By Solar System × Equipment Type: Helionis x Quantum, Oryn x Quantum, Flux x all system 

# SENSE CHECK HISTORICAL VS NB FOR INSIGHT
```{r}
# Compare new business vs historical for Helionis
cat("New business Helionis:\n")
new_business_ef %>%
  filter(solar_system == "Helionis Cluster") %>%
  summarise(
    avg_age     = round(mean(equipment_age),   2),
    avg_usage   = round(mean(usage_int),       2),
    avg_maint   = round(mean(maintenance_int), 0),
    avg_freq    = round(mean(pred_freq),       4)
  ) %>% print()

cat("\nHistorical Helionis:\n")
df_ef_freq_remapped %>%
  filter(solar_system == "Helionis Cluster") %>%
  summarise(
    avg_age     = round(mean(equipment_age),   2),
    avg_usage   = round(mean(usage_int),       2),
    avg_maint   = round(mean(maintenance_int), 0),
    avg_freq    = round(mean(claim_count/exposure), 4)
  ) %>% print()
```

# SIMULATION 
```{r}
# Monte Carlo Simulation 
set.seed(123)
n_sim      <- 100000
gamma_disp <- summary(sev_model_ef_final)$dispersion
gamma_shape <- 1 / gamma_disp

pred_freq_vec <- new_business_ef$pred_freq
pred_sev_vec  <- new_business_ef$pred_sev

# Simulation (no loading)
agg_loss_ef <- numeric(n_sim)

for (i in 1:n_sim) {
  claim_counts <- rpois(length(pred_freq_vec), lambda = pred_freq_vec)
  
  total_loss <- 0
  for (j in 1:length(pred_freq_vec)) {
    if (claim_counts[j] > 0) {
      gamma_rate <- gamma_shape / pred_sev_vec[j]
      claims     <- rgamma(claim_counts[j],
                           shape = gamma_shape,
                           rate  = gamma_rate)
      total_loss <- total_loss + sum(claims)
    }
  }
  agg_loss_ef[i] <- total_loss
}
```

# SIMULATION OUTPUT
```{r}
# Summary statistics 
cat("\n=== Aggregate Loss Distribution (No Loading) ===\n")
cat("Mean:          ", round(mean(agg_loss_ef),0), "\n")
cat("Std Deviation: ", round(sd(agg_loss_ef),0), "\n")
cat("Minimum:       ", round(min(agg_loss_ef),0), "\n")
cat("Maximum:       ", round(max(agg_loss_ef),0), "\n")

cat("\n=== Percentiles ===\n")
cat("50th (Median): ", round(quantile(agg_loss_ef, 0.50),0), "\n")
cat("75th:          ", round(quantile(agg_loss_ef, 0.75),0), "\n")
cat("90th:          ", round(quantile(agg_loss_ef, 0.90),0), "\n")
cat("95th:          ", round(quantile(agg_loss_ef, 0.95),0), "\n")
cat("99th (VaR):    ", round(quantile(agg_loss_ef, 0.99),0), "\n")
cat("99.95th (VaR):  ", round(quantile(agg_loss_ef, 0.995),0), "\n")

cat("\n=== Tail Risk (TVaR) ===\n")
var_95_ef  <- quantile(agg_loss_ef, 0.95)
var_99_ef  <- quantile(agg_loss_ef, 0.99)
var_995_ef <- quantile(agg_loss_ef, 0.995)
cat("TVaR 95%:  ", round(mean(agg_loss_ef[agg_loss_ef > var_95_ef]),0), "\n")
cat("TVaR 99%:  ", round(mean(agg_loss_ef[agg_loss_ef > var_99_ef]),0), "\n")
cat("TVaR 99.5%:", round(mean(agg_loss_ef[agg_loss_ef > var_995_ef]),0), "\n")
```

# SIMULATION VISUALIZATION
```{r}
# ── Plot ──────────────────────────────────────────────────────────────────────
ggplot(data.frame(agg_loss_ef), aes(x = agg_loss_ef)) +
  geom_histogram(bins = 100, fill = "steelblue", alpha = 0.7) +
  geom_vline(xintercept = var_95_ef, colour = "orange",
             linetype = "dashed", linewidth = 1) +
  geom_vline(xintercept = var_99_ef, colour = "red",
             linetype = "dashed", linewidth = 1) +
  annotate("text", x = var_95_ef, y = Inf,
           label = "VaR 95%", vjust = 2, colour = "orange") +
  annotate("text", x = var_99_ef, y = Inf,
           label = "VaR 99%", vjust = 2, colour = "red") +
  scale_x_continuous(labels = scales::comma) +
  labs(
    title = "Equipment Failure Aggregate Loss Distribution (No Loading)",
    x     = "Aggregate Loss",
    y     = "Count"
  ) +
  theme_minimal()
```

# CLOSE RESULT BETWEEN SIMULATION VS NON SIMULATION
```{r}
# compare 100000 simulation vs no simulation
sum(new_business_ef$pred_freq * new_business_ef$pred_sev) # no simulation
mean(agg_loss_ef) # simulation
```

# SIMULATION WITH LOADING  
```{r}
set.seed(123)
loading    <- 0.40
n_sim      <- 100000
gamma_disp <- summary(sev_model_ef_final)$dispersion

# Simulation
simulate_ef <- function(pred_freq_vec, pred_sev_vec, gamma_disp, loading, n_sim) {
  
  n_equipment <- length(pred_freq_vec)
  
  # Fixed earnings
  aggregate_earnings <- sum(pred_freq_vec * pred_sev_vec) * (1 + loading)
  
  results <- data.frame(
    aggregate_loss     = numeric(n_sim),
    aggregate_earnings = numeric(n_sim),
    profit             = numeric(n_sim),
    loss_ratio         = numeric(n_sim)
  )
  
  gamma_shape <- 1 / gamma_disp
  
  for (i in 1:n_sim) {
    claim_counts <- rpois(n_equipment, lambda = pred_freq_vec)
    
    total_loss <- 0
    for (j in 1:n_equipment) {
      if (claim_counts[j] > 0) {
        gamma_rate <- gamma_shape / pred_sev_vec[j]
        claims     <- rgamma(claim_counts[j],
                             shape = gamma_shape,
                             rate  = gamma_rate)
        total_loss <- total_loss + sum(claims)
      }
    }
    
    results$aggregate_loss[i]     <- total_loss
    results$aggregate_earnings[i] <- aggregate_earnings
    results$profit[i]             <- aggregate_earnings - total_loss
    results$loss_ratio[i]         <- total_loss / aggregate_earnings
  }
  results
}
```

# RUN SIMULATION
```{r}

cat("Running simulation...\n")
sim_results_ef <- simulate_ef(
  pred_freq_vec = new_business_ef$pred_freq,
  pred_sev_vec  = new_business_ef$pred_sev,
  gamma_disp    = gamma_disp,
  loading       = loading,
  n_sim         = n_sim
)
```

# SIMULATION OUTPUT SUMMARY
```{r}
cat("\n=== Aggregate Loss ===\n")
cat("Mean:        ", round(mean(sim_results_ef$aggregate_loss), 0), "\n")
cat("Median:      ", round(median(sim_results_ef$aggregate_loss), 0),  "\n")
cat("Std Dev:     ", round(sd(sim_results_ef$aggregate_loss),0), "\n")
cat("VaR 50%:     ", round(quantile(sim_results_ef$aggregate_loss,0.50),0), "\n")
cat("VaR 90%:     ", round(quantile(sim_results_ef$aggregate_loss,0.90),0), "\n")
cat("VaR 95%:     ", round(quantile(sim_results_ef$aggregate_loss, 0.95),0), "\n")
cat("VaR 99%:     ", round(quantile(sim_results_ef$aggregate_loss, 0.99),0), "\n")
cat("VaR 99.95%:   ", round(quantile(sim_results_ef$aggregate_loss, 0.9995), 0), "\n")
cat("TVaR 95%:    ", round(mean(sim_results_ef$aggregate_loss[sim_results_ef$aggregate_loss > quantile(sim_results_ef$aggregate_loss, 0.95)]),0), "\n")
cat("TVaR 99%:    ", round(mean(sim_results_ef$aggregate_loss[sim_results_ef$aggregate_loss > quantile(sim_results_ef$aggregate_loss, 0.99)]),0), "\n")
cat("TVaR 99.95%:  ", round(mean(sim_results_ef$aggregate_loss[sim_results_ef$aggregate_loss > quantile(sim_results_ef$aggregate_loss, 0.9995)]), 0), "\n")

cat("\n=== Earnings ===\n")
cat("Annual Premium (Earnings): ", round(mean(sim_results_ef$aggregate_earnings), 0), "\n")
cat("Median Premium:               ", round(median(sim_results_ef$aggregate_earnings), 0),  "\n")

cat("\n=== Profit ===\n")
cat("Mean Profit:        ", round(mean(sim_results_ef$profit), 0), "\n")
cat("Median Profit:      ", round(median(sim_results_ef$profit),0),  "\n")
cat("Std Dev:            ", round(sd(sim_results_ef$profit),0), "\n")
cat("Prob of Loss:       ", round(mean(sim_results_ef$profit < 0),4), "\n")
cat("VaR 50% (Profit):   ", round(quantile(sim_results_ef$profit, 0.50),0), "\n")
cat("VaR 90% (Profit):   ", round(quantile(sim_results_ef$profit, 0.10),0), "\n")
cat("VaR 95% (Profit):   ", round(quantile(sim_results_ef$profit, 0.05),0), "\n")
cat("VaR 99% (Profit):   ", round(quantile(sim_results_ef$profit, 0.01),0), "\n")
cat("VaR 99.95% (Profit): ", round(quantile(sim_results_ef$profit, 0.0005),0), "\n")

cat("\n=== Loss Ratio ===\n")
cat("Mean Loss Ratio:    ", round(mean(sim_results_ef$loss_ratio),4), "\n")
cat("Median Loss Ratio:  ", round(median(sim_results_ef$loss_ratio),4), "\n")
```

# SIMULATION RESULT Plots
```{r}
# Plots
par(mfrow = c(1, 3))

hist(sim_results_ef$aggregate_loss,
     breaks = 50,
     main   = "Aggregate Loss Distribution",
     xlab   = "Aggregate Loss",
     col    = "steelblue")

hist(sim_results_ef$profit,
     breaks = 50,
     main   = "Profit Distribution",
     xlab   = "Profit",
     col    = "seagreen")

hist(sim_results_ef$loss_ratio,
     breaks = 50,
     main   = "Loss Ratio Distribution",
     xlab   = "Loss Ratio",
     col    = "tomato")
```

# SENSITIVITY TESTING
```{r}
# Loading sensitivity 
set.seed(123)
loadings <- c(0.10, 0.20, 0.30, 0.40, 0.50)

loading_results_ef <- do.call(rbind, lapply(loadings, function(l) {
  
sim <- simulate_ef(
    pred_freq_vec = new_business_ef$pred_freq,
    pred_sev_vec  = new_business_ef$pred_sev,
    gamma_disp    = gamma_disp,
    loading       = l,
    n_sim         = 10000
  )
  
  data.frame(
    Loading        = paste0(l * 100, "%"),
    Mean_Earnings  = round(mean(sim$aggregate_earnings),0),
    Mean_Loss      = round(mean(sim$aggregate_loss), 0),
    Mean_Profit    = round(mean(sim$profit),0),
    Prob_of_Loss   = round(mean(sim$profit < 0), 4),
    Mean_LossRatio = round(mean(sim$loss_ratio), 4),
    VaR99_Loss     = round(quantile(sim$aggregate_loss, 0.99), 0),
    TVaR99_Loss    = round(mean(sim$aggregate_loss[sim$aggregate_loss > quantile(sim$aggregate_loss, 0.99)]), 0)
  )
}))

print(loading_results_ef)
```

#LONG TERM PROJECTION
# ECONOMIC ASSUMPTION
```{r}
avg_inflation <- 0.025   # 2.5%
avg_discount  <- 0.018   # 1.8%

cat("Inflation rate:", round(avg_inflation * 100, 2), "%\n")
cat("Discount rate:",  round(avg_discount  * 100, 2), "%\n")
```

# GROWTH ASSUMPTIONS - STABLE GROWTH
```{r}
base_helionis <- 2580
base_bayesia  <- 1290
base_oryn     <- 860

annual_growth_helionis <- (1 + 0.25)^(1/10) - 1
annual_growth_bayesia  <- (1 + 0.25)^(1/10) - 1
annual_growth_oryn     <- (1 + 0.15)^(1/10) - 1
```

# SENSE CHECK SOLARS AND EQUIPS
```{r}
equipment_levels <- freq_model_eff_final$xlevels$equipment_type
solar_levels     <- freq_model_eff_final$xlevels$solar_system

print(equipment_levels)
print(solar_levels)
```

# EQUIPS MIX PROPOTION BY SOLARS
```{r}
prop_helionis <- c(
  "Quantum Bore"       = 300/2580,
  "Graviton Extractor" = 240/2580,
  "FexStram Carrier"   = 150/2580,
  "ReglAggregators"    = 300/2580,
  "Flux Rider"         = 1500/2580,
  "Ion Pulverizer"     = 90/2580
)

prop_bayesia <- c(
  "Quantum Bore"       = 150/1290,
  "Graviton Extractor" = 120/1290,
  "FexStram Carrier"   = 75/1290,
  "ReglAggregators"    = 150/1290,
  "Flux Rider"         = 750/1290,
  "Ion Pulverizer"     = 45/1290
)

prop_oryn <- c(
  "Quantum Bore"       = 100/860,
  "Graviton Extractor" = 80/860,
  "FexStram Carrier"   = 50/860,
  "ReglAggregators"    = 100/860,
  "Flux Rider"         = 500/860,
  "Ion Pulverizer"     = 30/860
)
```

# BUILD SUPPORTING DATAFRAME
```{r}
age_dist_tbl <- data.frame(
  solar_system = c(
    rep(solar_levels[solar_levels %in% c("Helionis Cluster")], 6),
    rep(solar_levels[solar_levels %in% c("Bayesia System")], 6),
    rep(solar_levels[solar_levels %in% c("Oryn Delta")], 6)
  ),
  equipment_type = c(
    "Quantum Bore", "Graviton Extractor", "FexStram Carrier",
    "ReglAggregators", "Flux Rider", "Ion Pulverizer",
    "Quantum Bore", "Graviton Extractor", "FexStram Carrier",
    "ReglAggregators", "Flux Rider", "Ion Pulverizer",
    "Quantum Bore", "Graviton Extractor", "FexStram Carrier",
    "ReglAggregators", "Flux Rider", "Ion Pulverizer"
  ),
  lt5 = c(
    30/300, 24/240, 15/150, 30/300, 150/1500, 9/90,
    45/150, 36/120, 23/75, 45/150, 225/750, 14/45,
    75/100, 60/80, 37/50, 75/100, 375/500, 22/30
  ),
  yr5_9 = c(
    45/300, 36/240, 23/150, 45/300, 225/1500, 14/90,
    38/150, 30/120, 19/75, 38/150, 187/750, 11/45,
    15/100, 12/80, 8/50, 15/100, 75/500, 5/30
  ),
  yr10_14 = c(
    180/300, 144/240, 89/150, 180/300, 900/1500, 53/90,
    59/150, 48/120, 29/75, 59/150, 300/750, 18/45,
    10/100, 8/80, 5/50, 10/100, 50/500, 3/30
  ),
  yr15_19 = c(
    30/300, 24/240, 15/150, 30/300, 150/1500, 9/90,
    8/150, 6/120, 4/75, 8/150, 38/750, 2/45,
    0/100, 0/80, 0/50, 0/100, 0/500, 0/30
  ),
  yr20plus = c(
    15/300, 12/240, 8/150, 15/300, 75/1500, 5/90,
    0/150, 0/120, 0/75, 0/150, 0/750, 0/45,
    0/100, 0/80, 0/50, 0/100, 0/500, 0/30
  ),
  stringsAsFactors = FALSE
)

usage_tbl <- data.frame(
  solar_system = c(
    rep(solar_levels[solar_levels %in% c("Helionis Cluster")], 6),
    rep(solar_levels[solar_levels %in% c("Bayesia System")], 6),
    rep(solar_levels[solar_levels %in% c("Oryn Delta")], 6)
  ),
  equipment_type = rep(
    c("Quantum Bore", "Graviton Extractor", "FexStram Carrier",
      "ReglAggregators", "Flux Rider", "Ion Pulverizer"),
    3
  ),
  usage_pct = c(
    0.95, 0.95, 0.90, 0.80, 0.80, 0.50,
    0.80, 0.80, 0.75, 0.75, 0.80, 0.60,
    0.75, 0.75, 0.70, 0.70, 0.75, 0.50
  ),
  maintenance_int = c(
    750, 750, 375, 1500, 1500, 1000,
    600, 600, 400, 1000, 1000, 750,
    500, 500, 250, 300, 300, 500
  ),
  stringsAsFactors = FALSE
)

risk_tbl <- data.frame(
  solar_system = c(
    rep(solar_levels[solar_levels %in% c("Helionis Cluster")], 6),
    rep(solar_levels[solar_levels %in% c("Bayesia System")], 6),
    rep(solar_levels[solar_levels %in% c("Oryn Delta")], 6)
  ),
  equipment_type = rep(
    c("Quantum Bore", "Graviton Extractor", "FexStram Carrier",
      "ReglAggregators", "Flux Rider", "Ion Pulverizer"),
    3
  ),
  risk_idx = c(
    0.69, 0.48, 0.67, 0.24, 0.24, 0.64,
    0.77, 0.57, 0.71, 0.26, 0.21, 0.66,
    0.93, 0.73, 0.78, 0.35, 0.20, 0.75
  ),
  stringsAsFactors = FALSE
)
```

# LONG TERM NB GENERATOR
```{r}
make_nb_data_ef_lt <- function(
    n_helionis, n_bayesia, n_oryn,
    prop_helionis, prop_bayesia, prop_oryn,
    usage_tbl, risk_tbl, age_dist_tbl,
    equipment_levels, solar_levels
) {
  
  helionis_df <- data.frame(
    solar_system   = "Helionis Cluster",
    equipment_type = names(prop_helionis),
    n_equipment    = round(prop_helionis * n_helionis),
    stringsAsFactors = FALSE
  )
  
  bayesia_df <- data.frame(
    solar_system   = "Bayesia System",
    equipment_type = names(prop_bayesia),
    n_equipment    = round(prop_bayesia * n_bayesia),
    stringsAsFactors = FALSE
  )
  
  oryn_df <- data.frame(
    solar_system   = "Oryn Delta",
    equipment_type = names(prop_oryn),
    n_equipment    = round(prop_oryn * n_oryn),
    stringsAsFactors = FALSE
  )
  
  equipment_counts <- dplyr::bind_rows(helionis_df, bayesia_df, oryn_df)
  
  equipment_units <- tidyr::uncount(equipment_counts, weights = n_equipment)
  
  age_df <- equipment_units %>%
    dplyr::left_join(age_dist_tbl, by = c("solar_system", "equipment_type"))
  
  sampled_age_band <- apply(
    age_df[, c("lt5", "yr5_9", "yr10_14", "yr15_19", "yr20plus")],
    1,
    function(p) sample(c("lt5", "yr5_9", "yr10_14", "yr15_19", "yr20plus"),
                       size = 1, prob = p)
  )
  
  age_df$age_band <- sampled_age_band
  
  # treated as continuous intervals not integer year 
  age_df$equipment_age <- dplyr::case_when(
    age_df$age_band == "lt5"      ~ runif(nrow(age_df), 0, 5),
    age_df$age_band == "yr5_9"    ~ runif(nrow(age_df), 5, 10),
    age_df$age_band == "yr10_14"  ~ runif(nrow(age_df), 10, 15),
    age_df$age_band == "yr15_19"  ~ runif(nrow(age_df), 15, 20),
    age_df$age_band == "yr20plus" ~ runif(nrow(age_df), 20, 30)
  )
  
  nb_data <- age_df %>%
    dplyr::left_join(usage_tbl, by = c("solar_system", "equipment_type")) %>%
    dplyr::left_join(risk_tbl,  by = c("solar_system", "equipment_type")) %>%
    dplyr::mutate(
      exposure = 1,
      usage_int = usage_pct * 24
    ) %>%
    dplyr::select(
      equipment_type, solar_system, equipment_age,
      maintenance_int, usage_int, risk_idx, exposure
    )
  
  # Force exact model factor levels
  nb_data$equipment_type <- factor(nb_data$equipment_type, levels = equipment_levels)
  nb_data$solar_system   <- factor(nb_data$solar_system,   levels = solar_levels)
  
  nb_data
}
```
service-year categories were converted into continuous age values by sampling unifromly within each age interval.

# LT SIMULATION FUNCTION WITH LOADING 
```{r}
simulate_ef_segment <- function(pred_freq_vec, pred_sev_vec, gamma_disp, loading, n_sim) {
  
  n_units <- length(pred_freq_vec)
  gamma_shape <- 1 / gamma_disp
  
  premium_revenue <- sum(pred_freq_vec * pred_sev_vec) * (1 + loading)
  
  results <- data.frame(
    aggregate_cost = numeric(n_sim),
    revenue        = numeric(n_sim),
    net_revenue    = numeric(n_sim),
    loss_ratio     = numeric(n_sim)
  )
  
  for (i in seq_len(n_sim)) {
    claim_counts <- rpois(n_units, lambda = pred_freq_vec)
    
    total_loss <- 0
    for (j in seq_along(pred_freq_vec)) {
      if (claim_counts[j] > 0) {
        gamma_rate <- gamma_shape / pred_sev_vec[j]
        claims <- rgamma(claim_counts[j], shape = gamma_shape, rate = gamma_rate)
        total_loss <- total_loss + sum(claims)
      }
    }
    
    results$aggregate_cost[i] <- total_loss
    results$revenue[i]        <- premium_revenue
    results$net_revenue[i]    <- premium_revenue - total_loss
    results$loss_ratio[i]     <- total_loss / premium_revenue
  }
  
  results
}
```

# PARAMETERS
```{r}
set.seed(123)
n_sim   <- 100000    # each run = full 10 years
n_years <- 10
loading <- 0.40

gamma_disp <- summary(sev_model_ef_final)$dispersion
```

# PRE BUILD 10-YRS NB DATA
```{r}

cat("Building new business data for all 10 years...\n")

yearly_nb_data <- list()  # stores fleet data for each year

for (yr in 1:n_years) {
  
  cat("  Building year", yr, "data...\n")
  
  # Fleet size
  n_hel <- round(base_helionis * (1 + annual_growth_helionis)^yr)
  n_bay <- round(base_bayesia  * (1 + annual_growth_bayesia)^yr)
  n_ory <- round(base_oryn     * (1 + annual_growth_oryn)^yr)
  
  # Build data
  nb_data_yr <- make_nb_data_ef_lt(
    n_helionis       = n_hel,
    n_bayesia        = n_bay,
    n_oryn           = n_ory,
    prop_helionis    = prop_helionis,
    prop_bayesia     = prop_bayesia,
    prop_oryn        = prop_oryn,
    usage_tbl        = usage_tbl,
    risk_tbl         = risk_tbl,
    age_dist_tbl     = age_dist_tbl,
    equipment_levels = equipment_levels,
    solar_levels     = solar_levels
  )
  
  # Safety checks
  nb_data_yr$equipment_type <- factor(nb_data_yr$equipment_type,levels = equipment_levels)
  nb_data_yr$solar_system   <- factor(nb_data_yr$solar_system,levels = solar_levels)
  nb_data_yr <- nb_data_yr[!is.na(nb_data_yr$equipment_type) &!is.na(nb_data_yr$solar_system), ]
  
  # Predictions
  nb_data_yr$pred_freq <- predict(freq_model_eff_final,newdata = nb_data_yr,type    = "response")
  nb_data_yr$pred_sev  <- predict(sev_model_ef_final,newdata = nb_data_yr,type    = "response")
  
  # Apply inflation to severity
  inflation_factor    <- (1 + avg_inflation)^yr
  nb_data_yr$pred_sev <- nb_data_yr$pred_sev * inflation_factor
  
  # Calculate discount factor
  discount_factor <- 1 / (1 + avg_discount)^yr
  
  # Store
  yearly_nb_data[[yr]] <- list(
    nb_data         = nb_data_yr,
    n_exposures     = nrow(nb_data_yr),
    inflation_factor = inflation_factor,
    discount_factor  = discount_factor,
    
    # Pre-calculate annual earnings (fixed premium per year)
    annual_earnings  = sum(nb_data_yr$pred_freq * nb_data_yr$pred_sev) * (1 + loading)
  )
}

cat("All 10 years data built! Now running joint simulation...\n")

```

# JOINT 10-YRS SIMULATION - SIMULATE ALL 10-YRS EACH RUN THEN SUM UP
```{r}

# Storage for joint simulation results
joint_sim <- data.frame(
  # Nominal cumulative totals per run
  total_cost        = numeric(n_sim),  # sum of 10 years losses
  total_revenue     = numeric(n_sim),  # sum of 10 years premiums
  total_netrev      = numeric(n_sim),  # sum of 10 years profits
  
  # PV cumulative totals per run
  pv_total_cost     = numeric(n_sim),
  pv_total_revenue  = numeric(n_sim),
  pv_total_netrev   = numeric(n_sim)
)

cat("Running", n_sim, "joint simulations...\n")

for (i in 1:n_sim) {
  
  if (i %% 100 == 0) cat("  Simulation run", i, "of", n_sim, "...\n")
  
  run_total_cost       <- 0
  run_total_revenue    <- 0
  run_total_netrev     <- 0
  run_pv_total_cost    <- 0
  run_pv_total_revenue <- 0
  run_pv_total_netrev  <- 0
  
  # Loop through all 10 years within this single run
  for (yr in 1:n_years) {
    
    pred_freq_yr    <- yearly_nb_data[[yr]]$nb_data$pred_freq
    pred_sev_yr     <- yearly_nb_data[[yr]]$nb_data$pred_sev
    annual_earnings <- yearly_nb_data[[yr]]$annual_earnings
    discount_factor <- yearly_nb_data[[yr]]$discount_factor
    n_equipment_yr  <- yearly_nb_data[[yr]]$n_exposures
    
    # Simulate claims for this year
    claim_counts <- rpois(n_equipment_yr, lambda = pred_freq_yr)
    
    yr_total_loss <- 0
    for (j in 1:n_equipment_yr) {
      if (claim_counts[j] > 0) {
        gamma_shape <- 1 / gamma_disp
        gamma_rate  <- gamma_shape / pred_sev_yr[j]
        claims      <- rgamma(claim_counts[j],
                              shape = gamma_shape,
                              rate  = gamma_rate)
        yr_total_loss <- yr_total_loss + sum(claims)
      }
    }
    
    yr_netrev <- annual_earnings - yr_total_loss
    
    # Accumulate nominal
    run_total_cost    <- run_total_cost    + yr_total_loss
    run_total_revenue <- run_total_revenue + annual_earnings
    run_total_netrev  <- run_total_netrev  + yr_netrev
    
    # Accumulate PV
    run_pv_total_cost    <- run_pv_total_cost    + yr_total_loss    * discount_factor
    run_pv_total_revenue <- run_pv_total_revenue + annual_earnings  * discount_factor
    run_pv_total_netrev  <- run_pv_total_netrev  + yr_netrev        * discount_factor
  }
  
  # Store this run's 10-year totals
  joint_sim$total_cost[i]       <- run_total_cost
  joint_sim$total_revenue[i]    <- run_total_revenue
  joint_sim$total_netrev[i]     <- run_total_netrev
  joint_sim$pv_total_cost[i]    <- run_pv_total_cost
  joint_sim$pv_total_revenue[i] <- run_pv_total_revenue
  joint_sim$pv_total_netrev[i]  <- run_pv_total_netrev
}

cat("Joint simulation complete!\n")

```

# SENSE CHECK
```{r}
# Annual premium for Year 1
annual_yr1 <- yearly_nb_data[[1]]$annual_earnings
cat("Year 1 annual premium:", round(annual_yr1, 0), "\n")

# 10-year aggregate premium
cat("10-year aggregate premium:", round(mean(joint_sim$total_revenue), 0), "\n")

# Quick check: should be roughly annual × 10 (plus growth)
cat("10x Year 1:", round(annual_yr1 * 10, 0), "\n")
```

# SUMMARY OUTPUT
```{r}
# Risk count
avg_risk_count   <- round(mean(sapply(yearly_nb_data, function(x) x$n_exposures)), 0)
total_risk_count <- round(sum(sapply(yearly_nb_data, function(x) x$n_exposures)), 0)

cat("\n=== FOR EXCEL TABLE ===\n")
cat("Average Annual Risk Count:  ", avg_risk_count, "\n")
cat("Total Risk Count (10 yr):   ", total_risk_count, "\n")
cat("Loading:                     40%\n")

# NOMINAL 
cat("\n=== NOMINAL (Future Value) ===\n")

cat("\n--- Aggregate Premium ---\n")
cat("Total (Fixed):  ", round(mean(joint_sim$total_revenue),0), "\n")
cat("Mean:           ", round(mean(joint_sim$total_revenue),0), "\n")
cat("Median:         ", round(median(joint_sim$total_revenue),0), "\n")
cat("50%:            ", round(quantile(joint_sim$total_revenue, 0.50), 0), "\n")
cat("90%:            ", round(quantile(joint_sim$total_revenue, 0.90), 0), "\n")
cat("95%:            ", round(quantile(joint_sim$total_revenue, 0.95), 0), "\n")
cat("99%:            ", round(quantile(joint_sim$total_revenue, 0.99), 0), "\n")
cat("99.95%:          ", round(quantile(joint_sim$total_revenue, 0.9995),0), "\n")

cat("\n--- Aggregate Loss ---\n")
cat("Mean:           ", round(mean(joint_sim$total_cost), 0), "\n")
cat("Median:         ", round(median(joint_sim$total_cost), 0), "\n")
cat("StdDev:         ", round(sd(joint_sim$total_cost),0), "\n")
cat("50%:            ", round(quantile(joint_sim$total_cost, 0.50),0), "\n")
cat("90%:            ", round(quantile(joint_sim$total_cost, 0.90),0), "\n")
cat("95%:            ", round(quantile(joint_sim$total_cost, 0.95),0), "\n")
cat("99%:            ", round(quantile(joint_sim$total_cost, 0.99),0), "\n")
cat("99.95%:          ", round(quantile(joint_sim$total_cost, 0.9995),0), "\n")
cat("TVaR 99%:       ", round(mean(joint_sim$total_cost[
                        joint_sim$total_cost >
                        quantile(joint_sim$total_cost, 0.99)]),0), "\n")
cat("TVaR 99.95%:     ", round(mean(joint_sim$total_cost[
                        joint_sim$total_cost >
                        quantile(joint_sim$total_cost, 0.9995)]),0), "\n")

cat("\n--- Aggregate Profit (Net Revenue) ---\n")
cat("Mean:           ", round(mean(joint_sim$total_netrev),              0), "\n")
cat("Median:         ", round(median(joint_sim$total_netrev),            0), "\n")
cat("StdDev:         ", round(sd(joint_sim$total_netrev),                0), "\n")
cat("50%:            ", round(quantile(joint_sim$total_netrev, 0.50),    0), "\n")
cat("90%:            ", round(quantile(joint_sim$total_netrev, 0.10),    0), "\n")
cat("95%:            ", round(quantile(joint_sim$total_netrev, 0.05),    0), "\n")
cat("99%:            ", round(quantile(joint_sim$total_netrev, 0.01),    0), "\n")
cat("99.95%:          ", round(quantile(joint_sim$total_netrev, 0.0005),   0), "\n")
cat("Prob of Loss:   ", round(mean(joint_sim$total_netrev < 0), 4),       "\n")

# PRESENT VALUE 
cat("\n=== PRESENT VALUE (Today's Dollars) ===\n")

cat("\n--- PV Aggregate Premium ---\n")
cat("Total (Fixed):  ", round(mean(joint_sim$pv_total_revenue),           0), "\n")
cat("Mean:           ", round(mean(joint_sim$pv_total_revenue),           0), "\n")
cat("Median:         ", round(median(joint_sim$pv_total_revenue),         0), "\n")
cat("50%:            ", round(quantile(joint_sim$pv_total_revenue, 0.50), 0), "\n")
cat("90%:            ", round(quantile(joint_sim$pv_total_revenue, 0.90), 0), "\n")
cat("95%:            ", round(quantile(joint_sim$pv_total_revenue, 0.95), 0), "\n")
cat("99%:            ", round(quantile(joint_sim$pv_total_revenue, 0.99), 0), "\n")
cat("99.95%:          ", round(quantile(joint_sim$pv_total_revenue, 0.9995),0), "\n")

cat("\n--- PV Aggregate Loss ---\n")
cat("Mean:           ", round(mean(joint_sim$pv_total_cost),              0), "\n")
cat("Median:         ", round(median(joint_sim$pv_total_cost),            0), "\n")
cat("StdDev:         ", round(sd(joint_sim$pv_total_cost),                0), "\n")
cat("50%:            ", round(quantile(joint_sim$pv_total_cost, 0.50),    0), "\n")
cat("90%:            ", round(quantile(joint_sim$pv_total_cost, 0.90),    0), "\n")
cat("95%:            ", round(quantile(joint_sim$pv_total_cost, 0.95),    0), "\n")
cat("99%:            ", round(quantile(joint_sim$pv_total_cost, 0.99),    0), "\n")
cat("99.95%:          ", round(quantile(joint_sim$pv_total_cost, 0.9995),   0), "\n")
cat("TVaR 99%:       ", round(mean(joint_sim$pv_total_cost[
                        joint_sim$pv_total_cost >
                        quantile(joint_sim$pv_total_cost, 0.99)]),         0), "\n")
cat("TVaR 99.95%:     ", round(mean(joint_sim$pv_total_cost[
                        joint_sim$pv_total_cost >
                        quantile(joint_sim$pv_total_cost, 0.9995)]),        0), "\n")

cat("\n--- PV Aggregate Profit (Net Revenue) ---\n")
cat("Mean:           ", round(mean(joint_sim$pv_total_netrev),              0), "\n")
cat("Median:         ", round(median(joint_sim$pv_total_netrev),            0), "\n")
cat("StdDev:         ", round(sd(joint_sim$pv_total_netrev),                0), "\n")
cat("50%:            ", round(quantile(joint_sim$pv_total_netrev, 0.50),    0), "\n")
cat("90%:            ", round(quantile(joint_sim$pv_total_netrev, 0.10),    0), "\n")
cat("95%:            ", round(quantile(joint_sim$pv_total_netrev, 0.05),    0), "\n")
cat("99%:            ", round(quantile(joint_sim$pv_total_netrev, 0.01),    0), "\n")
cat("99.95%:          ", round(quantile(joint_sim$pv_total_netrev, 0.0005),   0), "\n")
cat("PV Prob of Loss:", round(mean(joint_sim$pv_total_netrev < 0), 4),       "\n")
```

########## STRESS TESTING
# BASE ASSUMPTIONS
```{r}
library(dplyr)
library(tidyr)

# 1. BASE ASSUMPTIONS
set.seed(123)

n_sim   <- 100000
loading <- 0.40

# Use fitted Gamma dispersion from final severity model
gamma_disp <- summary(sev_model_ef_final)$dispersion
```

# BASE PREDICTIONS AND FIXED PREMIUM
```{r}
# Base predicted outputs
base_pred_freq <- new_business_ef$pred_freq
base_pred_sev  <- new_business_ef$pred_sev

# Fixed base premium revenue for all stress scenarios
base_premium_revenue <- sum(base_pred_freq * base_pred_sev) * (1 + loading)

```

# SIMULATION FUNCTION WITH FIXED PREMIUM
```{r}
simulate_ef_fixed_premium <- function(pred_freq_vec, pred_sev_vec,
                                      gamma_disp, fixed_premium, n_sim) {
  
  n_equipment <- length(pred_freq_vec)
  gamma_shape <- 1 / gamma_disp
  
  results <- data.frame(
    aggregate_loss     = numeric(n_sim),
    aggregate_earnings = rep(fixed_premium, n_sim),
    profit             = numeric(n_sim),
    loss_ratio         = numeric(n_sim)
  )
  
  for (i in 1:n_sim) {
    claim_counts <- rpois(n_equipment, lambda = pred_freq_vec)
    
    total_loss <- 0
    for (j in 1:n_equipment) {
      if (claim_counts[j] > 0) {
        gamma_rate <- gamma_shape / pred_sev_vec[j]
        claims <- rgamma(claim_counts[j],
                         shape = gamma_shape,
                         rate  = gamma_rate)
        total_loss <- total_loss + sum(claims)
      }
    }
    
    results$aggregate_loss[i]     <- total_loss
    results$profit[i]             <- fixed_premium - total_loss
    results$loss_ratio[i]         <- total_loss / fixed_premium
  }
  
  results
}
```

# STRESS-TEST WRAPPER
```{r}
run_ef_stress_1yr <- function(freq_stress = 1.00,
                              sev_stress  = 1.00,
                              scenario_name = "Base") {
  
  stressed_freq <- base_pred_freq * freq_stress
  stressed_sev  <- base_pred_sev  * sev_stress
  
  sim_results <- simulate_ef_fixed_premium(
    pred_freq_vec = stressed_freq,
    pred_sev_vec  = stressed_sev,
    gamma_disp    = gamma_disp,
    fixed_premium = base_premium_revenue,
    n_sim         = n_sim
  )
  
  var50  <- as.numeric(quantile(sim_results$aggregate_loss, 0.50))
  var90  <- as.numeric(quantile(sim_results$aggregate_loss, 0.90))
  var95  <- as.numeric(quantile(sim_results$aggregate_loss, 0.95))
  var99  <- as.numeric(quantile(sim_results$aggregate_loss, 0.99))
  var995 <- as.numeric(quantile(sim_results$aggregate_loss, 0.9995))
  
  tvar95  <- mean(sim_results$aggregate_loss[sim_results$aggregate_loss > var95])
  tvar99  <- mean(sim_results$aggregate_loss[sim_results$aggregate_loss > var99])
  tvar995 <- mean(sim_results$aggregate_loss[sim_results$aggregate_loss > var995])
  
  profit50  <- as.numeric(quantile(sim_results$profit, 0.50))   # ← new
  profit90  <- as.numeric(quantile(sim_results$profit, 0.10))   # ← new lower tail
  profit95  <- as.numeric(quantile(sim_results$profit, 0.05))   # ← new lower tail
  profit99  <- as.numeric(quantile(sim_results$profit, 0.01))   # ← new lower tail
  profit995 <- as.numeric(quantile(sim_results$profit, 0.0005))  # ← new lower tail
  
  
  data.frame(
    Scenario              = scenario_name,
    N_Exposures           = length(base_pred_freq),
    Fixed_Premium_Revenue = base_premium_revenue,
    
    Mean_Aggregate_Loss   = mean(sim_results$aggregate_loss),
    SD_Aggregate_Loss     = sd(sim_results$aggregate_loss),
    
    VaR_50_Loss           = var50,   
    VaR_90_Loss           = var90,   
    VaR_95_Loss           = var95,
    VaR_99_Loss           = var99,
    VaR_99_5_Loss         = var995,
    
    # TVaR
    TVaR_95_Loss          = tvar95,
    TVaR_99_Loss          = tvar99,
    TVaR_99_5_Loss        = tvar995,
    # Profit statistics
    Mean_Net_Revenue      = mean(sim_results$profit),
    SD_Net_Revenue        = sd(sim_results$profit),
    # Profit percentiles
    Profit_50             = profit50,   # ← new
    Profit_90             = profit90,   # ← new
    Profit_95             = profit95,   # ← new
    Profit_99             = profit99,   # ← new
    Profit_99_5           = profit995,  # ← new
    # Risk metrics
    Prob_Loss             = mean(sim_results$profit < 0),
    Mean_Loss_Ratio       = mean(sim_results$loss_ratio)
  )
}
```

# RUN SCENARIOS
```{r}
ef_st_base <- run_ef_stress_1yr(
  freq_stress   = 1.00,
  sev_stress    = 1.00,
  scenario_name = "Base"
)

ef_st_freq <- run_ef_stress_1yr(
  freq_stress   = 1.20,
  sev_stress    = 1.00,
  scenario_name = "Frequency +20%"
)

ef_st_sev <- run_ef_stress_1yr(
  freq_stress   = 1.00,
  sev_stress    = 1.20,
  scenario_name = "Severity +20%"
)

ef_st_combined <- run_ef_stress_1yr(
  freq_stress   = 1.20,
  sev_stress    = 1.20,
  scenario_name = "Freq +20% & Sev +20%"
)

ef_stress_1yr_summary <- dplyr::bind_rows(
  ef_st_base,
  ef_st_freq,
  ef_st_sev,
  ef_st_combined
)
```

# RESULT 
```{r}
ef_stress_1yr_summary %>%
  mutate(
    across(
      c(
        N_Exposures, Fixed_Premium_Revenue, Mean_Aggregate_Loss,
        SD_Aggregate_Loss, Mean_Net_Revenue, SD_Net_Revenue,
        VaR_95_Loss, VaR_99_Loss, VaR_99_5_Loss,
        TVaR_95_Loss, TVaR_99_Loss, TVaR_99_5_Loss,
        Profit_50, Profit_90, Profit_95, Profit_99, Profit_99_5
      ),
      ~ round(.x, 0)
    ),
    Mean_Loss_Ratio = round(Mean_Loss_Ratio, 4),
    Prob_Loss       = round(Prob_Loss, 4)
  ) %>%
  print()
```


# FORMATTED OUTPUT
```{r}
cat("\n=== STRESS TEST RESULTS — SHORT TERM (1 YEAR) ===\n")
cat("Base Premium Revenue:", round(base_premium_revenue, 0), "\n")
cat("Number of Exposures: ", nrow(base_pred_freq), "\n")
cat("Loading:              40%\n\n")

# Full summary table 
cat("=== FULL SUMMARY TABLE ===\n")
ef_stress_1yr_summary %>%
  mutate(across(where(is.numeric), ~ round(.x, 0))) %>%
  as.data.frame() %>%
  print()

# FOR EXCEL TABLE ANALYSIS
cat("\n=== FOR EXCEL TABLE ===\n")

for (i in 1:nrow(ef_stress_1yr_summary)) {
  row <- ef_stress_1yr_summary[i, ]
  cat("\n--- Scenario:", row$Scenario, "---\n")
  cat("Risk Count:       ", row$N_Exposures,                          "\n")
  cat("Loading:           40%\n")
 cat("\n--- Aggregate Premium ---\n")
  cat("Total (Fixed):", round(row$Fixed_Premium_Revenue,    0), "\n")
  
  cat("\n--- Aggregate Loss ---\n")
  cat("Mean:         ", round(row$Mean_Aggregate_Loss,      0), "\n")
  cat("Std Dev:      ", round(row$SD_Aggregate_Loss,        0), "\n")
  cat("Median:       ", round(row$VaR_50_Loss,              0), "\n")
  cat("50%:          ", round(row$VaR_50_Loss,              0), "\n")
  cat("90%:          ", round(row$VaR_90_Loss,              0), "\n")
  cat("95%:          ", round(row$VaR_95_Loss,              0), "\n")
  cat("99%:          ", round(row$VaR_99_Loss,              0), "\n")
  cat("99.5%:        ", round(row$VaR_99_5_Loss,            0), "\n")
  cat("TVaR 95%:     ", round(row$TVaR_95_Loss,             0), "\n")
  cat("TVaR 99%:     ", round(row$TVaR_99_Loss,             0), "\n")
  cat("TVaR 99.5%:   ", round(row$TVaR_99_5_Loss,           0), "\n")
  
  cat("\n--- Aggregate Profit (Net Revenue) ---\n")
  cat("Mean:         ", round(row$Mean_Net_Revenue,         0), "\n")
  cat("Std Dev:      ", round(row$SD_Net_Revenue,           0), "\n")
  cat("Median:       ", round(row$Profit_50,                0), "\n")
  cat("50%:          ", round(row$Profit_50,                0), "\n")
  cat("90%:          ", round(row$Profit_90,                0), "\n")
  cat("95%:          ", round(row$Profit_95,                0), "\n")
  cat("99%:          ", round(row$Profit_99,                0), "\n")
  cat("99.5%:        ", round(row$Profit_99_5,              0), "\n")
  
  cat("\n--- Risk Metrics ---\n")
  cat("Prob of Loss: ", round(row$Prob_Loss * 100,          4), "%\n")
  cat("Loss Ratio:   ", round(row$Mean_Loss_Ratio * 100,    2), "%\n")
  
  cat("\n--- Capital Gap above Premium ---\n")
  cap_gap_99  <- max(0, row$VaR_99_Loss   - row$Fixed_Premium_Revenue)
  cap_gap_995 <- max(0, row$VaR_99_5_Loss - row$Fixed_Premium_Revenue)
  cat("At VaR 99%:   ", round(cap_gap_99,                   0), "\n")
  cat("At VaR 99.5%: ", round(cap_gap_995,                  0), "\n")
}
```

# COMPARE TABLE 
```{r}
# Comparison table (change vs base) 
cat("\n=== CHANGE VS BASE SCENARIO ===\n")
base_loss <- ef_stress_1yr_summary$Mean_Aggregate_Loss[1]
base_profit <- ef_stress_1yr_summary$Mean_Net_Revenue[1]
base_var99 <- ef_stress_1yr_summary$VaR_99_Loss[1]

ef_stress_1yr_summary %>%
  mutate(
    Change_Mean_Loss   = round(Mean_Aggregate_Loss - base_loss,0),
    Pct_Change_Loss    = round((Mean_Aggregate_Loss/base_loss - 1) * 100, 2),
    Change_Mean_Profit = round(Mean_Net_Revenue - base_profit,0),
    Change_VaR99       = round(VaR_99_Loss - base_var99,0),
    Capital_Gap_995    = round(pmax(0, VaR_99_5_Loss - Fixed_Premium_Revenue), 0)
  ) %>%
  dplyr::select(
    Scenario,
    Mean_Aggregate_Loss,
    Change_Mean_Loss,
    Pct_Change_Loss,
    Mean_Net_Revenue,
    Change_Mean_Profit,
    VaR_99_Loss,
    Change_VaR99,
    Prob_Loss,
    Mean_Loss_Ratio,
    Capital_Gap_995
  ) %>%
  mutate(across(where(is.numeric), ~ round(.x, 0))) %>%
  as.data.frame() %>% 
  print()  
```


# SCENARIO TESTING RDS EXTRACT
```{r}
saveRDS(
  sim_results_ef$aggregate_loss,
  file = paste0(bronze_path, "ef_base_agg_loss.rds")
)

cat("Done! Length:", length(sim_results_ef$aggregate_loss), "\n")
# Should print: 100000
```

# RDS FULL DATA FOR REFERENCE 
```{r}
# Save FULL simulation results for EQF
saveRDS(
  sim_results_ef,
  file = paste0(bronze_path, "ef_short_term_sim_full_results.rds")
)

cat("Full EQF results saved!\n")
cat("Rows:", nrow(sim_results_ef), "\n")
cat("Columns:", ncol(sim_results_ef), "\n")
cat("Column names:", names(sim_results_ef), "\n")
```



